---
title: "Which College is Best for You? Part 2"
author: "Michael L. Thompson"
date: "January 17, 2016, rev. Oct. 25, 2017"
output:
  html_document:
    toc: yes
    toc_depth: 4
  pdf_document:
    toc: yes
    toc_depth: '4'
always_allow_html: yes
---
# Introduction
This is a followup to my first script ["Which College is Best for You?_"](https://www.kaggle.com/apollostar/d/kaggle/college-scorecard/which-college-is-best-for-you).

Ben Hamner suggested in his comments to that first script that I write a script allowing anyone to specify his or her student profile right at the top. In this way the script might be more appealing for other folks to fork and play with it themselves.  Good advice!

So here it is.

One thing though,... Using RStudio on this R Markdown file, you can click on "Chunks-->Run All" to save the variables in the global environment. Then it won't have to re-process the database every time you make a change to the student profile and re-run it each time with "Chunks-->Run All" to generate a new top schools list.

## By the Way...

I've now added an R Shiny GUI and made it into a browser-based app: ["Best Colleges for You"](https://thompsonml.shinyapps.io/BestCollegeApp/)!


# Specify Your Student Profile
```{r setup,echo=FALSE, message=FALSE, warning=FALSE}
library('RSQLite')   # for SQLite manipulation of database
library('magrittr')  # for piping with infix %>%, %<>%, %T>% (for cool, sophistication effect)
library('tidyverse')
library('gridExtra') # for grid layout of ggplots
library('leaflet')   # for mapping of the top schools for each student
library('htmltools') # for assistance with leaflet
library('xtable')    # for rendering tables in HTML
```

```{r profile}
Case <- list(studentProfile = list(
  dependent = TRUE,           # c(TRUE,FALSE)
  ethnicity = 'white',        # c('white','black','hispanic','asian')
  gender    = 'male',       # c('male','female')
  age       = 'le24',         # c('le24','gt24') (years of age)
  income    = 'gt30Kle110K',       # c('le30K','gt30Kle110K','gt110K') ($US)
  earnings  = 'gt110K',       # c('le30K','gt30Kle75K','gt75Kle110K','gt110K') ($US)
  sat       = 'gt1000le1200',       # c('le800','gt800le1000','gt1000le1200','gt1200le1400','gt1400')
  fafsa     = 'fsend_3',      # c('fsend_1','fsend_2','fsend_3','fsend_4','fsend_5')
  discipline= 'ComputerInformation',  # See below.
  region    = 'GreatLakes',      # See below for which states in which regions.
  locale    = 'CityMidsize',    # See below.
  traits    = list(           # See below.
    Risk      = 'M',
    Vision    = 'M',
    Breadth   = 'M',
    Challenge = 'M') # Must be in order; only 'L','M','H' valid.
)
)
```

(Note: 
The comments show the only valid values for each field.
The `fafsa` field refers to how many schools FAFSA form was sent to. Values `'fsend_1'` to `'fsend_4'` correspond to 1 to 4 schools, resp., and `'fsend_5'` refers to 5 or more schools.
Values for `income`, `earnings` and `sat` contain `"le"` meaning "less than or equal to" and `"gt"` meaning "greater than". The `"K"` in the `income` and `earnings` values means "thousand".  So `'gt30Kle110K'` means "greater than $30,000 and less than or equal to $110,00".

The fields with "See below." are explained in section ***"Explanation of Valid Entries for Student Profile"*** at the end of the file.)


# Setup the Data & Model
The next chunk loads the college database and defines necessary data structures and functions to implement the model.


```{r makeDSFunc, results='hide', echo=FALSE, message=FALSE, warning=FALSE}
#================================================================
# Function make_DataSpec: Loads the college database, processes into Bayes factors, and defines the data structures need for modeling.
make_DataSpec <- function(){
  # Academic disciplines:
  discgrp <- data_frame(LABEL=c("Agriculture, Agriculture Operations, and Related Sciences",
                                "Natural Resources and Conservation",
                                "Architecture and Related Services",
                                "Area, Ethnic, Cultural, Gender, and Group Studies",
                                "Communication, Journalism, and Related Programs",
                                "Communications Technologies/Technicians and Support Services",
                                "Computer and Information Sciences and Support Services",
                                "Personal and Culinary Services",
                                "Education",
                                "Engineering",
                                "Engineering Technologies and Engineering-Related Fields",
                                "Foreign Languages, Literatures, and Linguistics",
                                "Family and Consumer Sciences/Human Sciences",
                                "Legal Professions and Studies",
                                "English Language and Literature/Letters",
                                "Liberal Arts and Sciences, General Studies and Humanities",
                                "Library Science",
                                "Biological and Biomedical Sciences",
                                "Mathematics and Statistics",
                                "Military Technologies and Applied Sciences",
                                "Multi/Interdisciplinary Studies",
                                "Parks, Recreation, Leisure, and Fitness Studies",
                                "Philosophy and Religious Studies",
                                "Theology and Religious Vocations",
                                "Physical Sciences",
                                "Science Technologies/Technicians","Psychology",
                                "Homeland Security, Law Enforcement, Firefighting and Related Protective Services",
                                "Public Administration and Social Service Professions","Social Sciences",
                                "Construction Trades",
                                "Mechanic and Repair Technologies/Technicians",
                                "Precision Production",
                                "Transportation and Materials Moving",
                                "Visual and Performing Arts",
                                "Health Professions and Related Programs",
                                "Business, Management, Marketing, and Related Support Services",
                                "History"),
                        discgrp=c(2,4,2,5,6,4,1,4,5,1,1,5,5,7,5,5,5,2,1,4,5,4,5,5,2,3,3,4,3,3,4,4,4,4,6,2,7,5),
                        VARIABLE.NAME=c("PCIP01","PCIP03","PCIP04","PCIP05","PCIP09","PCIP10","PCIP11","PCIP12","PCIP13",
                                        "PCIP14","PCIP15","PCIP16","PCIP19","PCIP22","PCIP23","PCIP24","PCIP25","PCIP26",
                                        "PCIP27","PCIP29","PCIP30","PCIP31","PCIP38","PCIP39","PCIP40","PCIP41","PCIP42",
                                        "PCIP43","PCIP44","PCIP45","PCIP46","PCIP47","PCIP48","PCIP49","PCIP50","PCIP51",
                                        "PCIP52","PCIP54"))
  
  discNames <- gsub('^([A-Z][-a-z]+)[, and/]*([A-Z][a-z]+).*','\\1\\2',discgrp$LABEL)
  discgrp %<>% tbl_df %>% mutate(discName = discNames) 
  
  # Get the college database: if not already exist, data wrangle & compute Bayes factors. 
  cat("Loading college database ...")
  #  if(!exists('studentBF') | !exists('student')){
  #     if(file.exists("studentBF.RData")){
  #       load("studentBF.RData")
  #     } 
  #     if(file.exists("student.RData")){
  #       load("student.RData")
  #     }
  #    if(!exists('studentBF') | !exists('student')){
  # If still don't have these objects 'student' & 'studentBF', build them from the database file.
  
  db <- dbConnect(dbDriver("SQLite"), "../input/database.sqlite")
  # This stops SQLite from writing temp files to disk, which use all the available space
  dbGetQuery(db, "PRAGMA temp_store=2;") 
  
  fieldNames <-
    c(
      'UGDS','UGDS_WHITE','UGDS_BLACK','UGDS_HISP','UGDS_ASIAN','UGDS_AIAN','UGDS_NHPI',
      'UGDS_2MOR','UGDS_NRA','UGDS_UNKN',"UG25abv","INC_PCT_LO","DEP_STAT_PCT_IND","DEP_INC_PCT_LO","IND_INC_PCT_LO",
      "DEP_INC_PCT_M1","DEP_INC_PCT_M2","DEP_INC_PCT_H1","IND_INC_PCT_M1","IND_INC_PCT_M2","IND_INC_PCT_H1","IND_INC_PCT_H2",
      "DEP_INC_PCT_H2","PAR_ED_PCT_1STGEN","pct_white","pct_black","pct_asian","pct_hispanic","pct_ba","pct_grad_prof",
      "pct_born_us","md_faminc","median_hh_inc","poverty_rate","unemp_rate","ln_median_hh_inc","pell_ever","female","fsend_1","fsend_2",
      "fsend_3","fsend_4","fsend_5","INC_PCT_LO","INC_PCT_M1","INC_PCT_M2","INC_PCT_H1","INC_PCT_H2","LATITUDE",
      "LONGITUDE","ZIP","STABBR","region","LOCALE","CCSIZSET","DISTANCEONLY","RELAFFIL","CURROPER","NPT4_PUB",
      "NPT4_PRIV","NPT41_PUB","NPT42_PUB","NPT43_PUB","NPT44_PUB","NPT45_PUB","NPT41_PRIV","NPT42_PRIV","NPT43_PRIV",
      "NPT44_PRIV","NPT45_PRIV","TUITIONFEE_IN","TUITIONFEE_OUT","PCTPELL","ADM_RATE","RET_FT4","SATVR25","SATVR75",
      "SATMT25","SATMT75","SATWR25","SATWR75","SATVRMID","SATMTMID","SATWRMID","ACTCM25","ACTCM75","ACTEN25","ACTEN75",
      "ACTMT25","ACTMT75","ACTWR25","ACTWR75","ACTCMMID","ACTENMID","ACTMTMID","ACTWRMID","SAT_AVG","SAT_AVG_ALL",
      "PCIP01","PCIP03","PCIP04","PCIP05","PCIP09","PCIP10","PCIP11","PCIP12",
      "PCIP13","PCIP14","PCIP15","PCIP16","PCIP19","PCIP22","PCIP23","PCIP24","PCIP25","PCIP26","PCIP27","PCIP29","PCIP30",
      "PCIP31","PCIP38","PCIP39","PCIP40","PCIP41","PCIP42","PCIP43","PCIP44","PCIP45","PCIP46","PCIP47","PCIP48","PCIP49",
      "PCIP50","PCIP51","PCIP52","PCIP54","C150_4_WHITE","C150_4_BLACK","C150_4_HISP","C150_4_ASIAN","ENRL_ORIG_YR2_RT",
      "LO_INC_ENRL_ORIG_YR2_RT","MD_INC_ENRL_ORIG_YR2_RT","HI_INC_ENRL_ORIG_YR2_RT","DEP_ENRL_ORIG_YR2_RT","IND_ENRL_ORIG_YR2_RT",
      "FEMALE_ENRL_ORIG_YR2_RT","MALE_ENRL_ORIG_YR2_RT","PELL_ENRL_ORIG_YR2_RT","NOPELL_ENRL_ORIG_YR2_RT","LOAN_ENRL_ORIG_YR2_RT",
      "NOLOAN_ENRL_ORIG_YR2_RT","FIRSTGEN_ENRL_ORIG_YR2_RT","NOT1STGEN_ENRL_ORIG_YR2_RT","ENRL_ORIG_YR3_RT","LO_INC_ENRL_ORIG_YR3_RT",
      "MD_INC_ENRL_ORIG_YR3_RT","HI_INC_ENRL_ORIG_YR3_RT","DEP_ENRL_ORIG_YR3_RT","IND_ENRL_ORIG_YR3_RT","FEMALE_ENRL_ORIG_YR3_RT",
      "MALE_ENRL_ORIG_YR3_RT","PELL_ENRL_ORIG_YR3_RT","NOPELL_ENRL_ORIG_YR3_RT","LOAN_ENRL_ORIG_YR3_RT","NOLOAN_ENRL_ORIG_YR3_RT",
      "FIRSTGEN_ENRL_ORIG_YR3_RT","NOT1STGEN_ENRL_ORIG_YR3_RT","ENRL_ORIG_YR4_RT","LO_INC_ENRL_ORIG_YR4_RT","MD_INC_ENRL_ORIG_YR4_RT",
      "HI_INC_ENRL_ORIG_YR4_RT","DEP_ENRL_ORIG_YR4_RT","IND_ENRL_ORIG_YR4_RT","FEMALE_ENRL_ORIG_YR4_RT","MALE_ENRL_ORIG_YR4_RT",
      "PELL_ENRL_ORIG_YR4_RT","NOPELL_ENRL_ORIG_YR4_RT","LOAN_ENRL_ORIG_YR4_RT","NOLOAN_ENRL_ORIG_YR4_RT","FIRSTGEN_ENRL_ORIG_YR4_RT",
      "NOT1STGEN_ENRL_ORIG_YR4_RT","ENRL_ORIG_YR6_RT","LO_INC_ENRL_ORIG_YR6_RT","MD_INC_ENRL_ORIG_YR6_RT","HI_INC_ENRL_ORIG_YR6_RT",
      "DEP_ENRL_ORIG_YR6_RT","IND_ENRL_ORIG_YR6_RT","FEMALE_ENRL_ORIG_YR6_RT","MALE_ENRL_ORIG_YR6_RT","PELL_ENRL_ORIG_YR6_RT",
      "NOPELL_ENRL_ORIG_YR6_RT","LOAN_ENRL_ORIG_YR6_RT","NOLOAN_ENRL_ORIG_YR6_RT","FIRSTGEN_ENRL_ORIG_YR6_RT","NOT1STGEN_ENRL_ORIG_YR6_RT",
      "ENRL_ORIG_YR8_RT","LO_INC_ENRL_ORIG_YR8_RT","MD_INC_ENRL_ORIG_YR8_RT","HI_INC_ENRL_ORIG_YR8_RT","DEP_ENRL_ORIG_YR8_RT",
      "IND_ENRL_ORIG_YR8_RT","FEMALE_ENRL_ORIG_YR8_RT","MALE_ENRL_ORIG_YR8_RT","PELL_ENRL_ORIG_YR8_RT","NOPELL_ENRL_ORIG_YR8_RT",
      "LOAN_ENRL_ORIG_YR8_RT","NOLOAN_ENRL_ORIG_YR8_RT","FIRSTGEN_ENRL_ORIG_YR8_RT","NOT1STGEN_ENRL_ORIG_YR8_RT","C150_4_POOLED_SUPP",
      "PCTFLOAN","DEBT_MDN","GRAD_DEBT_MDN","WDRAW_DEBT_MDN","LO_INC_DEBT_MDN","MD_INC_DEBT_MDN","HI_INC_DEBT_MDN","DEP_DEBT_MDN",
      "IND_DEBT_MDN","PELL_DEBT_MDN","NOPELL_DEBT_MDN","CUML_DEBT_N","CUML_DEBT_P90","CUML_DEBT_P75","CUML_DEBT_P25",
      "CUML_DEBT_P10","loan_ever","DEBT_MDN_SUPP","GRAD_DEBT_MDN_SUPP","GRAD_DEBT_MDN10YR_SUPP","CDR2","CDR3","COMPL_RPY_1YR_RT",
      "NONCOM_RPY_1YR_RT","LO_INC_RPY_1YR_RT","MD_INC_RPY_1YR_RT","HI_INC_RPY_1YR_RT","DEP_RPY_1YR_RT","IND_RPY_1YR_RT",
      "PELL_RPY_1YR_RT","NOPELL_RPY_1YR_RT","FEMALE_RPY_1YR_RT","MALE_RPY_1YR_RT","FIRSTGEN_RPY_1YR_RT","NOTFIRSTGEN_RPY_1YR_RT",
      "RPY_3YR_RT","COMPL_RPY_3YR_RT","NONCOM_RPY_3YR_RT","LO_INC_RPY_3YR_RT","MD_INC_RPY_3YR_RT","HI_INC_RPY_3YR_RT","DEP_RPY_3YR_RT",
      "IND_RPY_3YR_RT","PELL_RPY_3YR_RT","NOPELL_RPY_3YR_RT","FEMALE_RPY_3YR_RT","MALE_RPY_3YR_RT","FIRSTGEN_RPY_3YR_RT",
      "NOTFIRSTGEN_RPY_3YR_RT","RPY_5YR_RT","COMPL_RPY_5YR_RT","NONCOM_RPY_5YR_RT","LO_INC_RPY_5YR_RT","MD_INC_RPY_5YR_RT",
      "HI_INC_RPY_5YR_RT","DEP_RPY_5YR_RT","IND_RPY_5YR_RT","PELL_RPY_5YR_RT","NOPELL_RPY_5YR_RT","FEMALE_RPY_5YR_RT",
      "MALE_RPY_5YR_RT","FIRSTGEN_RPY_5YR_RT","NOTFIRSTGEN_RPY_5YR_RT","RPY_7YR_RT","COMPL_RPY_7YR_RT","NONCOM_RPY_7YR_RT",
      "LO_INC_RPY_7YR_RT","MD_INC_RPY_7YR_RT","HI_INC_RPY_7YR_RT","DEP_RPY_7YR_RT","IND_RPY_7YR_RT","PELL_RPY_7YR_RT",
      "NOPELL_RPY_7YR_RT","FEMALE_RPY_7YR_RT","MALE_RPY_7YR_RT","FIRSTGEN_RPY_7YR_RT","NOTFIRSTGEN_RPY_7YR_RT","count_ed",
      "count_nwne_p6","count_wne_p6","mn_earn_wne_p6","md_earn_wne_p6","pct10_earn_wne_p6","pct25_earn_wne_p6",
      "pct75_earn_wne_p6","pct90_earn_wne_p6","sd_earn_wne_p6","gt_25k_p6","mn_earn_wne_inc1_p6","mn_earn_wne_inc2_p6",
      "mn_earn_wne_inc3_p6"
    )
  # This dropped name set is part of the base query string...
  fieldNames <- setdiff(fieldNames,c('CITY','LONGITUDE','LATITUDE','ZIP','region','LOCALE','CURROPER'))
  # drop the discipline fields; they're grabbed separately. 
  fieldNames <- grep('PCIP',fieldNames,value = TRUE,invert = TRUE) 
  
  # Upper case variables are from 2013 and lower case variables are from the
  # Treasury dataset of 2005, except 'region'. fromS10 <-
  # grep('ENRL_ORIG',fieldNames)
  fromS11 <- which(fieldNames %in% grep('^[A-Z]|region',fieldNames))
  fromS05 <- setdiff(seq_along(fieldNames),fromS11)
  # put a '_2005' suffix on the Treasury variables.
  dfNames <- ifelse(grepl('^[A-Z]|region',fieldNames),fieldNames,paste(fieldNames,"2005",sep="_")) 
  
  makeQuery <- function(year,fieldNames,dfNames) {
    paste("SELECT UNITID unitID,
          INSTNM College,
          CONTROL CollegeType,
          PREDDEG degree,
          CURROPER currop,
          DISTANCEONLY distance,
          RELAFFIL relaffil,
          CITY city,
          LONGITUDE lon,
          LATITUDE  lat,
          ZIP zip,
          st_fips state,
          region region,
          LOCALE locale,
          CCBASIC ccbasic,
          CCUGPROF ccugprof,
          Year Year,",
          paste(fieldNames,' ',dfNames,sep="",collapse = ","),
          "FROM Scorecard 
          WHERE Year=",year)
  }
  
  queryString2005 <- makeQuery(2005,fieldNames,dfNames)
  queryString2013 <- makeQuery(2013,fieldNames,dfNames)
  
  
  #       discNames <- gsub('^([A-Z][-a-z]+)[, and/]*([A-Z][a-z]+).*','\\1\\2',discgrp$LABEL)
  #       discgrp %<>% tbl_df %>% mutate(discName = discNames)
  queryStringDscplns <- makeQuery(2013,discgrp$VARIABLE.NAME,discNames)
  
  # Retrieve the data for 2005 because that is the latest with Treasury data on student families.
  student2005 <- dbGetQuery(db,queryString2005)
  # Retrieve the data for 2013 because that's the latest data.
  student2013 <- dbGetQuery(db,queryString2013)
  disciplines2013 <- dbGetQuery(db,queryStringDscplns)
  
  #       cat("\n")
  #       disciplines2013 %>% 
  #         filter(grepl('Harvard|Cal.+Inst.+Tech|Mass.+Inst.+Tech|Princeton U|Northwestern U|Cornell U',College)) %>% 
  #         dplyr::select(College,Engineering) %>%
  #         mutate(EngineeringProb = 100*Engineering) %>%
  #         show
  
  
  # Disconnect the database.
  dbDisconnect(db)
  
  #*** 2016/06/11: Discovered that Princeton University has "0" for "ComputerInformation"!
  disciplines2013 %>% 
    filter( College == "Princeton University" ) %>% 
    select( 1:2, ComputerInformation, Engineering ) %>% 
    print
  # Based on this web page (https://www.cs.princeton.edu/people/ugrad), I estimated ~450 computer science majors 
  # out of the 5234 undergraduates enrolled at Princeton.# And since the Computer Science dept. is in the 
  # Engineering school, I'll set "ComputerInformation" to 450/5234=8.6% and subtract that same amount from "Engineering".
  csprop <- 0.086
  engg   <-  disciplines2013[ disciplines2013$College == "Princeton University", "Engineering" ][[ 1 ]]
  disciplines2013[ disciplines2013$College == "Princeton University", c( "ComputerInformation", "Engineering") ] <- c( csprop, engg - csprop )
  disciplines2013 %>% 
    filter( College == "Princeton University" ) %>% 
    select( 1:2, ComputerInformation, Engineering ) %>% 
    print
  
  
  # Join the years together.
  student2013 <- student2013[!sapply(student2013,function(x) all(is.na(x)))] %>% tbl_df
  student2005 <- student2005[!sapply(student2005,function(x) all(is.na(x)))] %>% tbl_df
  # student <- student2013 %>% tbl_df %>% 
  #   dplyr::select(-contains('_2005')) %>% 
  #   inner_join(student2005 %>% dplyr::select(unitID,contains('_2005')), by = "unitID")
  student <- student2013 %>% 
    inner_join(student2005 %>% dplyr::select(unitID,contains('_2005')), by = "unitID")
  # Add the dominant disciplines
  discplns <- disciplines2013 %>% tbl_df %>%
    filter(unitID %in% student$unitID) %>%
    mutate(irow = seq_len(nrow(.))) %>%
    gather(key=Discipline,value=Proportion,one_of(discNames)) %>%
    group_by(irow) %>%
    arrange(desc(Proportion)) %>%
    summarize(
      unitID   = first(unitID),
      College  = first(College),
      domDisc1 = first(Discipline),
      pctDisc1 = 100*first(Proportion),
      domDisc2 = nth(Discipline,n=2),
      pctDisc2 = 100*nth(Proportion,n=2)
    ) %>%
    mutate(
      isSpecialty = pctDisc1 > 70,
      isTheological = grepl('Theolo|Relig',domDisc1) | grepl('Theolo|Relig',domDisc2)
    )
  
  # CALLED THE MAIN DATA TABLE `student` BECAUSE ORIGINALLY ONLY CONTAINED
  # `student` CATEGORY OF DATA FIELDS. NOW THIS LEGACY NAME AWKWARDLY CONFUSES
  # THIS DATA TABLE OF UNIVERSITIES/COLLEGES WITH THE OTHER DATA OBJECTS
  # CONTAINING SPECIFIC STUDENT PROPERTIES -- SEE LATER...
  
  # Now add the disciplines as grouped
  student %<>% 
    inner_join(discplns %>% dplyr::select(-College),by='unitID')
  
  usa.states.dc <- c("District of Columbia",state.name)
  # Filter down to just the colleges meeting the following criteria:
  student %<>% 
    filter(CollegeType != 'Private for-profit',
           currop      == 'Currently certified as operating',
           distance    == 'Not distance-education only',
           degree      == "Predominantly bachelor's-degree granting",
           as.character(region)      != 'U.S. Service Schools',
           !grepl('^Associate',ccbasic),
           state %in% usa.states.dc,
           !is.na(ccbasic),
           !is.na(UGDS),
           !is.na(pell_ever_2005),
           !isTheological,
           !is.na(mn_earn_wne_inc2_p6_2005))
  
  # Turn character columns into factors.
  student <- student %>% lapply(
    function(x) 
      if(is.character(x) | is.factor(x) | is.logical(x)) {
        if('PrivacySuppressed' %in% x) {
          tmp<-x;tmp[x=='PrivacySuppressed']<-NA;as.numeric(tmp)
        } else {
          factor(gsub("’","'",x))
        }
      } else as.numeric(x)) %>% 
    as.data.frame %>% tbl_df
  
  # add a religious affiliation indicator variable:
  student %<>% 
    mutate(relaffilFLAG=factor(ifelse(is.na(relaffil),
                                      FALSE,
                                      !grepl('^[Nn]ot ',as.character(relaffil)))))
  
  
  a<- student %>% dplyr::select(grep('^(SAT|ACT)',names(.))) %>%  apply(1,function(x) !all(is.na(x)))
  student %<>% mutate(stdzdtest=a)
  
  #show(dim(student))
  #show(summary(student))
  
  
  #       Drop all the Penn St. and U. of Pitt. schools except the main campus because
  #       same 2005 Treasury data entered in all rows for each, resp.
  
  stdttmp <- student %>% filter((College %in% c('Pennsylvania State University-Main Campus',
                                                'University of Pittsburgh-Pittsburgh Campus')) |
                                  !(College %in% grep('Pennsylvania State University|University of Pittsburgh',
                                                      College,value=TRUE)))
  #       show(setdiff(student$College,stdttmp$College))
  #       show(grep('Pennsylvania State University|University of Pittsburgh',stdttmp$College,value=TRUE))
  student <- stdttmp
  rm(stdttmp)
  student <- student[!sapply(student,function(x) length(which(is.na(x)))>1000)]
  # Drop schools without earnings data:
  #       student %>% filter(is.na(pct10_earn_wne_p6_2005) | pct10_earn_wne_p6_2005==0) %>% 
  #         dplyr::select(unitID,College,UGDS,matches('pct.+earn')) 
  student %<>% 
    filter(!is.na(pct10_earn_wne_p6_2005) & pct10_earn_wne_p6_2005>0) 
  
  ### Ethnicity:
  ethnicity <- student %>% 
    dplyr::select(unitID,College,contains("UGDS")) %>%
    mutate(probSchool = UGDS/sum(UGDS),
           totprob    = rowSums(as.matrix(.[4:ncol(.)])))
  #ethnicity %>% print(n=20)
  
  # This computes P(E | not(H)) the conditional probability of evidence given the
  # hypothesis is FALSE using inputs of cpH = P(E | H) probability of evidence
  # given the hypothesis is TRUE and pH = P(H) the marginal (prior) probability
  # of hypothesis being TRUE. But, in this case of almost 1500 schools, where "H"
  # is school si, P(E|not(H)) is approximately P(E) = sum(P(E|H)*P(H);over H). 
  # So, could skip this function and just use P(E).
  # 
  # ***IMPORTANT:*** I use log10 of the Bayes factors, which is a bit unusual,
  # but I know powers of 10 better than powers of 2 or *e*.
  
  cpNotH <- function(cpH,pH){
    pEH <- cpH * pH
    pE  <- sum(pEH)
    return((pE - pEH)/(1 - pH))
  }
  BFactor <- function(cpH,cpNotH,logFunc=log10){
    x <- cpH / cpNotH
    return(if(is.function(logFunc)) logFunc(x) else x)
  }
  makeCpNotH <- function(x) mapply(cpNotH,cpH=x,pH=ethnicity$probSchool)
  makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*ethnicity$probSchool,na.rm=TRUE))} # Approximate
  
  ethnicBF <- ethnicity %>% 
    dplyr::select(contains("UGDS_")) %>% 
    mutate_all(funs(makeBF)) %>% 
    setNames(gsub("UGDS","BF",names(.))) %$%
    bind_cols(ethnicity[1:2],.) %>%
    mutate(BF_female_2005 = makeBF(student$female_2005))
  #=====================================================
  
  ### Income: THESE VALUES ARE UNRELIABLE BECAUSE SO MANY MAJOR SCHOOLS HAVE MISSING OR ZERO VALUES.
  ###         USE DISTRIBUTIONS DERIVED FROM median_hh_inc_2005 INSTEAD.
  ###         ALSO ADD IN pct_born_us_2005 (for student origins) and loan_ever_2005 (for campus setting)
  
  income <- student %>% 
    dplyr::select(unitID,College,UGDS,DEP_STAT_PCT_IND,contains("INC_PCT")) %>%
    mutate(probSchool = UGDS/sum(UGDS),
           totprob      = rowSums(as.matrix(.[c("INC_PCT_LO","INC_PCT_M1","INC_PCT_M2","INC_PCT_H1",
                                                "INC_PCT_H2")])),
           totprobdep   = rowSums(as.matrix(.[c("DEP_INC_PCT_LO","DEP_INC_PCT_M1","DEP_INC_PCT_M2",
                                                "DEP_INC_PCT_H1","DEP_INC_PCT_H2")])),
           totprobind   = rowSums(as.matrix(.[c("IND_INC_PCT_LO","IND_INC_PCT_M1","IND_INC_PCT_M2",
                                                "IND_INC_PCT_H1","IND_INC_PCT_H2")])),
           totprob2dep  = rowSums(as.matrix(.[c("DEP_INC_PCT_LO","DEP_INC_PCT_H2")])),
           totprob2ind  = rowSums(as.matrix(.[c("IND_INC_PCT_LO","IND_INC_PCT_H2")])))
  #income %>% print(n=20)
  
  # To avoid dropping many schools due to missing values will hold 7 variables
  # for income: DEP_STAT_PCT_IND, DEP_INC_PCT_LO, DEP_INC_PCT_H2 &
  # DEP_INC_GTLO_LTH2, IND_INC_PCT_LO, IND_INC_PCT_H2 & IND_INC_GTLO_LTH2; where
  # the x_INC_GTLO_LTH2 variables are 1 - x_INC_PCT_H2 - x_INC_PCT_LO; and the
  # sum of the latter 6 columns = 1. So must mutate the current IND_... columns
  # by multiplying by DEP_STAT_PCT_IND. An "x" is appended to the end of the
  # newly computed variable names.
  
  income %<>% mutate(probSchool = UGDS/sum(UGDS),
                     DEP_STAT_PCT_INDx  = DEP_STAT_PCT_IND,
                     DEP_STAT_PCT_DEPx  = 1 - DEP_STAT_PCT_INDx,
                     DEP_INC_PCT_LOx    = DEP_INC_PCT_LO*DEP_STAT_PCT_DEPx,
                     DEP_INC_PCT_HI2x   = DEP_INC_PCT_H2*DEP_STAT_PCT_DEPx,
                     DEP_INC_GTLO_LTH2x = DEP_STAT_PCT_DEPx - DEP_INC_PCT_LOx - DEP_INC_PCT_HI2x,
                     IND_INC_PCT_LOx    = IND_INC_PCT_LO*DEP_STAT_PCT_INDx,
                     IND_INC_PCT_HI2x   = IND_INC_PCT_H2*DEP_STAT_PCT_INDx,
                     IND_INC_GTLO_LTH2x = DEP_STAT_PCT_IND - IND_INC_PCT_LOx - IND_INC_PCT_HI2x,
                     totprobx = DEP_INC_PCT_LOx + DEP_INC_PCT_HI2x + DEP_INC_GTLO_LTH2x + IND_INC_PCT_LOx
                     + IND_INC_PCT_HI2x + IND_INC_GTLO_LTH2x) %>%
    dplyr::select(unitID,College,UGDS,probSchool,matches("x$"))
  #income %>% print(n=20)
  #income[income$College=='California Institute of Technology',
  #        c('IND_INC_PCT_LOx','IND_INC_PCT_HI2x','IND_INC_GTLO_LTH2x','totprobx')] <- c(0,0,0,1)
  
  # Fit a lognormal to quantiles
  # Clumsily fit log-normal distributions to the upper & lower income quintiles for each school.
  fr <- function(x,INC_50,INC_low,INC_high,probs=c(0.20,0.80)) {
    sdl <- x[1]
    #q <- qlnorm(p=probs,meanlog=log(INC_MEAN) - sdl^2/2,sdlog=sdl)
    q <- qlnorm(p=probs,meanlog=log(INC_50),sdlog=sdl)
    log(INC_low/q[1])^2 + log(INC_high/q[2])^2
  }
  getMuSd <- function(INC_50,INC_low,INC_high,probs=c(0.20,0.80),fr = fr) {
    
    INC_500 <- INC_50
    
    soln <- optim(c(1), fr,lower=1.0E-3,upper=2,method='L-BFGS-B',
                  INC_50=INC_50,INC_low=INC_low,INC_high=INC_high,probs=probs)
    sdl  <- soln$par
    # meanlog  <- log(INC_MEAN) - sdl^2/2
    meanlog  <- log(INC_50)
    # pINC_MEAN <- exp(meanlog+0.5*sdl^2)
    pINC_50 <- exp(meanlog)
    q   <- qlnorm(p=probs,meanlog=meanlog,sdlog=sdl)
    
    return(list(meanlog=meanlog,sdlog=sdl,pINC_low=q[1],pINC_50=pINC_50,pINC_high=q[2],convergence=soln$convergence,message=soln$message))
  }      
  
  # Given the percentage of students in each income quintile, we fit the mean and std. dev. of the log-normal distribution
  # of student incomes.
  a <- student %>% 
    dplyr::select(unitID,College,contains('DEP_INC_'),md_faminc_2005,median_hh_inc_2005) %>% 
    filter(!is.na(DEP_INC_PCT_LO) & !is.na(DEP_INC_PCT_H2)) %>% 
    filter(DEP_INC_PCT_LO>0 & DEP_INC_PCT_H2 > 0)
  a %<>% 
    rowwise %>% 
    mutate(sdlog = getMuSd(md_faminc_2005,30000,110000,c(DEP_INC_PCT_LO,1.0-DEP_INC_PCT_H2-1.0E-12),fr=fr)[['sdlog']]) %T>% 
    print(width=Inf)
  sdlogEst   <- a %>% 
    dplyr::select(unitID,sdlog) %>% 
    right_join(student %>% dplyr::select(unitID,College,md_faminc_2005) ,by='unitID')
  sdlogLoess <- loess(sdlog ~ log(md_faminc_2005),data=a)
  sdlogPred  <- predict(sdlogLoess,newdata=sdlogEst)
  sdlogPred  <- data_frame(sdlogp = as.numeric(ifelse(is.na(sdlogEst$sdlog),sdlogPred,sdlogEst$sdlog)))
  sdlogEst %<>% bind_cols(sdlogPred)
  sdlogEst %<>%
    mutate(p_le30K  = plnorm( 30000,meanlog = log(md_faminc_2005),sdlog = sdlogp,lower.tail =  TRUE),
           p_gt110K = plnorm(110000,meanlog = log(md_faminc_2005),sdlog = sdlogp,lower.tail = FALSE),
           p_gt30Kle110K = 1 - p_le30K - p_gt110K
    )
  sdlogEst %<>% bind_cols(income %>% dplyr::select(probSchool)) %>% dplyr::select(unitID,College,probSchool,p_le30K,p_gt30Kle110K,p_gt110K) 
  makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*sdlogEst$probSchool,na.rm=TRUE))} # Approximate
  bf     <- sdlogEst %>% mutate_at(funs(makeBF),.vars=vars(starts_with('p_'))) %>% setNames(gsub('^p_','BF_',names(.)))
  
  # Check-in on some select schools:
  selectSchools <- paste(c('Massachusetts Institute of Technology', 'California Institute of Technology',
                           'Princeton', 'Yale', 'Harvard', 'Stanford', 'Duke', 'Vanderbilt', 'Berkeley',
                           'Northwestern U', 'Princeton', 'Cornell U', 'Ohio State U','University of Dayton'),collapse="|")
  sdlogEst %>% filter(grepl(selectSchools,College))
  
  
  makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*income$probSchool,na.rm=TRUE))} # Approximate
  incomeBF <- income %>% 
    dplyr::select(matches("x$"),-totprobx) %>% 
    mutate_all(funs(makeBF)) %>% 
    setNames(gsub("(.+)x$","BF_\\1x",names(.))) %$%
    bind_cols(income[1:4],.) %>% inner_join(bf %>% select(unitID,starts_with('BF_')),by='unitID')
  
  #===================================================
  
  ### Student Aid:
  
  aid <- student %>% 
    dplyr::select(unitID,College,UGDS,matches("pell_ever|fsend")) %>%
    mutate(probSchool = UGDS/sum(UGDS),
           totprob    = rowSums(as.matrix(.[grepl('fsend',names(.))])))
  #aid %>% print(n=20)
  makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*aid$probSchool,na.rm=TRUE))} # Approximate
  aidBF <- aid %>% 
    dplyr::select(matches("pell_ever|fsend")) %>% 
    mutate_all(funs(makeBF)) %>% 
    setNames(paste0("BF_",names(.))) %$%
    bind_cols(aid[1:3],.)
  
  #       g <- aidBF %>% gather(key=Variable,value=BF,-(1:3)) %>%
  #         ggplot(aes(x=BF,fill=Variable))
  #       g + geom_density(alpha=0.3) + 
  #         facet_wrap(~Variable) + 
  #         scale_x_continuous(lim=c(-1,1)) + 
  #         theme(text=element_text(size=14,face = 'bold'))
  #===================================================
  
  ### Earnings after graduation (6 years on)
  
  earnRepay <- student %>% 
    dplyr::select(unitID,College,UGDS,matches("^(RPY|CDR|.+earn_wne_p6)")) %>%
    mutate(probSchool = UGDS/sum(UGDS),
           CDR3est       = (CDR3*UGDS + 10)/(UGDS + 10 + 140), # smooth with 150 hypothetical students having a nominal default rate of ~7%
           sdlog         = sqrt(log((sd_earn_wne_p6_2005/mn_earn_wne_p6_2005)^2+1)),
           meanlog       = log(mn_earn_wne_p6_2005^2/sqrt(mn_earn_wne_p6_2005^2+sd_earn_wne_p6_2005^2)),
           p_le30K       = plnorm(30.0E3,meanlog,sdlog),
           p_gt30Kle48K  = plnorm(48.0E3,meanlog,sdlog)  - p_le30K,
           p_gt48Kle75K  = plnorm(75.0E3,meanlog,sdlog)  - plnorm(48.0E3,meanlog,sdlog),
           p_gt75Kle110K = plnorm(110.0E3,meanlog,sdlog) - plnorm(75.0E3,meanlog,sdlog),
           p_gt110K      = plnorm(110.0E3,meanlog,sdlog,lower.tail=FALSE),
           totprob       = p_le30K + p_gt30Kle48K + p_gt48Kle75K + p_gt75Kle110K + p_gt110K) %>%
    dplyr::select(1:3,CDR3,CDR3est,5:7,probSchool,starts_with('p_')) 
  
  makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*earnRepay$probSchool,na.rm=TRUE))} # Approximate
  earnRepayBF <- earnRepay %>% 
    dplyr::select(-c(1:3,9)) %>% 
    mutate_all(funs(makeBF)) %>% 
    setNames(paste0("BF_",names(.))) %$%
    bind_cols(earnRepay[c(1:3,9)],.) %>%
    filter(complete.cases(.))
  
  #       earnRepayBF %>% 
  #         dplyr::select(-3) %>% 
  #         inner_join(student %>% dplyr::select(unitID,ADM_RATE),by='unitID') %>% 
  #         filter(!is.na(ADM_RATE)) %>% 
  #         mutate(ADM_RATEd=discretize(.['ADM_RATE'],method='interval',breaks=6)[['ADM_RATE']]) %>% 
  #         arrange(desc(BF_p_gt110K)) %>% 
  #         filter(BF_p_le30K<0.5 & BF_p_gt110K>0.5) %$% 
  #         qplot(x=BF_p_le30K,y=BF_p_gt110K,color=ADM_RATEd) + geom_text(aes(label=College))
  #===================================================
  
  ### College Setting: locale & region.
  
  #       student %$% { locale %>% summary %>% print }
  #       student %$% { region %>% summary %>% print }
  
  localeNames <- sort(unique(as.character(student$locale)))
  localeAggregates <- setNames(c(gsub('([^ ]+) ([^ ]+).+','\\1\\2',localeNames[1:3]),
                                 rep('Rural',3),
                                 gsub('([^ ]+) ([^ ]+).+','\\1\\2',localeNames[7]),
                                 rep('Suburb:Small/Midsize & Town:Fringe',2),
                                 gsub('([^ ]+) ([^ ]+).+','\\1\\2',localeNames[10]),
                                 'Suburb:Small/Midsize & Town:Fringe',
                                 gsub('([^ ]+) ([^ ]+).+','\\1\\2',localeNames[12])),localeNames)
  student %<>% mutate(localeAgg = factor(localeAggregates[as.character(locale)],
                                         levels=c('Rural','Town:Remote','Town:Distant','Suburb:Small/Midsize & Town:Fringe',
                                                  'Suburb:Large','City:Small','City:Midsize','City:Large')))
  makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*settingBF$probSchool,na.rm=TRUE))} # Approximate
  settingBF <- student %>% 
    dplyr::select(unitID,College,UGDS,localeAgg,region) %>%
    mutate(probSchool = UGDS/sum(UGDS))
  a <- settingBF %$% model.matrix(~localeAgg - 1,data=.)
  localeColNames <- gsub('^([^:]+):*(.+)$','BF_\\1\\2',colnames(a))
  colnames(a) <- localeColNames
  locAbbr <- gsub('^.+Agg(.{4}).+','\\1',colnames(a))
  
  # Define similarity factors
  fctr <- matrix(c(1  ,0.1, 0.01,0.0001,
                   0.1,  1, 0.5 ,0.1,
                   0.01,0.5,  1, 0.6,
                   0.0001,0.1,0.6,  1)/2,nrow=4,ncol=4,byrow=TRUE, 
                 dimnames=list(key=c('Rura','Town','Subu','City'), other=c('Rura','Town','Subu','City')))
  # Expand factors to cover all locales.
  fctr <- fctr[locAbbr,locAbbr]
  fctr[row(fctr) == col(fctr)] <- 1 # put ones on the diagonal
  dimnames(fctr) <- list(localeColNames,localeColNames)
  
  # Reweight the local indicators.
  a <- a %*% fctr
  
  #       f <- apply(a,1,function(x,zone) {
  #         tmp <- x + fctr[zone[x==1],zone]
  #         return(tmp/sum(tmp))
  #         },zone=gsub('^BF_localeAgg(.{4}).+','\\1',colnames(a)))
  #       
  #       a[] <- c(t(f))
  
  # Convert locale indicators intor psuedo-Bayes factors.
  a %<>% as.data.frame %>% tbl_df %>% mutate_all(funs(makeBF))
  settingBF %<>% dplyr::select(-localeAgg) %>% bind_cols(a)
  
  #regNames <- c('FarWest','GreatLakes','MidEast','NewEngland','Plains','RockyMountains','Southeast','Southwest')
  a <- settingBF %>%
    mutate(regionName = factor(gsub(' ','',as.character(student$region)))) %$% 
    model.matrix(~regionName - 1,data=.)
  # Define similarity factors
  fctrRegion <- matrix(c(1, 0.03, 0.0, 0.0, 0.1, 0.3, 0.0, 0.3,
                         0.03, 1, 0.3, 0.1, 0.3, 0.1, 0.03, 0.03,
                         0.0, 0.3, 1,  0.3, 0.01, 0.0, 0.1, 0.0,
                         0.0, 0.1, 0.3, 1, 0.0, 0.0, 0.0, 0.0,
                         0.1, 0.3, 0.01, 0.0, 1, 0.3, 0.1, 0.1,
                         0.3, 0.1, 0.0, 0.0, 0.3, 1, 0.0, 0.3,
                         0.0, 0.03, 0.1, 0.0, 0.1, 0.0, 1, 0.3,
                         0.3,0.03, 0.0, 0.0, 0.1, 0.3, 0.3, 1),nrow=8,ncol=8,byrow=TRUE,
                       dimnames=list(colnames(a),colnames(a)))
  a <- a %*% fctrRegion
  colnames(a) <- gsub('^regionName','BF_',colnames(a))
  a %<>% as.data.frame %>% tbl_df %>% mutate_all(funs(makeBF))
  settingBF %<>% dplyr::select(-region) %>% bind_cols(a)
  
  # Didn't get around to adding in costs...
  #===================================================
  #costBF
  #===================================================
  ### Academics: Completion rates, admissions rates and SAT scores.
  # Took a look at scaled test scores, but ultimately settled upon the log-normal
  # fits to SATs below.
  # scaleSAT <- function(Score) (Score - 280)/(800 - 280)
  # scaleACT <- function(Score) (Score -   9)/( 36 -   9)
  # Impute missing values with the means
  #lmtest <- lm(ACTCMMID ~ SAT_AVG,data=student)
  
  # academicsBF <- student %>% 
  #   filter(!(isSpecialty=='TRUE' & is.na(ADM_RATE))) %>%
  #   dplyr::select(unitID,College,ADM_RATE,matches('(^C150)|(SAT|ACT)')) %>%
  #   filter(!is.na(C150_4_POOLED_SUPP)) %>%
  #   mutate(ADM_RATE = ifelse(is.na(ADM_RATE),mean(ADM_RATE,na.rm=TRUE),ADM_RATE)) %>%
  #   mutate(SAT_AVG_scaled  = scaleSAT(ifelse(is.na(SAT_AVG) , mean(SAT_AVG,na.rm=TRUE), SAT_AVG))) %>%
  #   mutate(ACTCMMID_scaled = scaleACT(ifelse(is.na(ACTCMMID),mean(ACTCMMID,na.rm=TRUE),ACTCMMID))) %>%
  #   dplyr::select(1:3,ncol(.) - (2:0))
  
  # a <- academicsBF %>% 
  #   mutate_each(funs(scaleSAT),starts_with('SAT'),-ends_with('scaled')) %>% 
  #   mutate_each(funs(scaleACT),starts_with('ACT'),-ends_with('scaled')) %>%
  #   gather(key=Test,value=scaledScore,-unitID,-College,-ADM_RATE)
  
  #***NOTE:*** This is a bit of a kludge, but it results in a distribution for SAT
  #Verbal+Math scores for each school, from which Bayes factors for 200-pt
  #intervals are computed. (Kludginess: Adding the quartiles of Verbal & Math
  #DOES NOT equal the quartiles of Verbal+Math.)
  
  student %>% 
    mutate(SAT_25=SATVR25+SATMT25,SAT_75=SATVR75+SATMT75) %>% 
    filter(!is.na(SAT_AVG) & !is.na(SAT_75)) %>% 
    ggplot(aes(x=SAT_AVG)) + geom_density(fill='orange',alpha=0.3) + 
    geom_density(aes(x=SAT_25),fill='red',alpha=0.1) + 
    geom_density(aes(x=SAT_75),fill='green',alpha=0.1)
  
  academics <- student %>% 
    mutate(SAT_25=SATVR25+SATMT25,SAT_75=SATVR75+SATMT75) %>% 
    filter(!is.na(SAT_AVG) & !is.na(SAT_75)) %>% 
    dplyr::select(unitID,College,UGDS,ADM_RATE,C150_4_POOLED_SUPP,SAT_25,SAT_AVG,SAT_75)
  
  
  # Clumsily fit log-normal distributions to the SAT score quartiles for each school.
  frSAT <- function(x,SAT_AVG,SAT_25,SAT_75,probs=c(0.25,0.75)) {
    sdl <- x[1]
    q <- qlnorm(p=probs,meanlog=log(SAT_AVG) - sdl^2/2,sdlog=sdl)
    log(SAT_25/q[1])^2 + log(SAT_75/q[2])^2
  }
  getMuSdSAT <- function(SAT_AVG,SAT_25,SAT_75,fr=frSAT) {
    probs  <- c(0.25,0.75)
    SAT_AVG0 <- SAT_AVG
    if(SAT_75==1600) {
      probs   <- probs/(0.75+1.0E-5)
      SAT_AVG <- (SAT_AVG0 - 1600*(1-(0.75+1.0E-5)))/(0.75+1.0E-5)
    }
    soln <- optim(c(0.05), fr,lower=1.0E-3,upper=0.2,method='L-BFGS-B',
                  SAT_AVG=SAT_AVG,SAT_25=SAT_25,SAT_75=SAT_75,probs=probs)
    sdl  <- soln$par
    meanlog  <- log(SAT_AVG) - sdl^2/2
    pSAT_AVG <- exp(meanlog+0.5*sdl^2)
    q   <- qlnorm(p=probs,meanlog=meanlog,sdlog=sdl)
    if(SAT_75==1600){
      pSAT_AVG <- pSAT_AVG*(0.75+1.0E-5) + 1600*(1-(0.75+1.0E-5))
    }
    return(c(meanlog=meanlog,sdlog=sdl,pSAT_25=q[1],pSAT_AVG=pSAT_AVG,pSAT_75=q[2]))
  }
  
  # Get the log-normal parameters of the SAT distribution for each school
  academics <- academics %$% {bind_cols(.,as.data.frame(t(mapply(getMuSdSAT,SAT_AVG,SAT_25,SAT_75,MoreArgs = list(fr=frSAT)))))}
  academics %>% 
    filter(grepl('Harvard|Cal.+Inst.+Tech|Mass.+Inst.+Tech|Princeton U|Northwestern U|Cornell U',College)) %>% 
    arrange(desc(SAT_AVG))
  
  # Discretize the SAT distributions
  academics %<>%
    filter(!is.na(C150_4_POOLED_SUPP)) %>%
    mutate(probSchool = UGDS/sum(UGDS),
           p_le800        = plnorm( 800,meanlog,sdlog),
           p_gt800le1000  = plnorm(1000,meanlog,sdlog) - p_le800,
           p_gt1000le1200 = plnorm(1200,meanlog,sdlog) - plnorm(1000,meanlog,sdlog),
           p_gt1200le1400 = plnorm(1400,meanlog,sdlog) - plnorm(1200,meanlog,sdlog),
           p_gt1400       = plnorm(1400,meanlog,sdlog,lower.tail=FALSE),
           totprob        = p_le800 + p_gt800le1000 + p_gt1000le1200 + p_gt1200le1400 + p_gt1400)
  
  # Compute Bayes Factors for the discretized SAT score distributions of each school
  makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*academics$probSchool,na.rm=TRUE))} # Approximate
  academicsBF <- academics %>% 
    dplyr::select(-c(1:3,matches('SAT|log|prob'))) %>% 
    mutate_all(funs(makeBF)) %>% 
    setNames(gsub("^p","BF_SAT",names(.))) %>%
    setNames(gsub("^([^B])","BF_\\1",names(.))) %$%
    bind_cols(academics[1:3],.)
  
  # Now add in the disciplines:
  makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*academics$probSchool,na.rm=TRUE))} # Approximate
  discplnsBF <-  academics %>% 
    dplyr::select(unitID,probSchool) %>% 
    inner_join(disciplines2013,by='unitID') %>%
    mutate_at(funs(makeBF),.vars=vars(-seq_len(which(names(.) == discNames[1])-1))) %>% 
    setNames(c(setdiff(names(.),discNames),paste0('BF_',discNames)))
  
  
  
  academicsBF %<>% inner_join(discplnsBF %>% dplyr::select(unitID,starts_with('BF_')),by='unitID')
  b <- disciplines2013 %>% filter(sapply(unitID,function(ui) ui %in% academicsBF$unitID)) %>% dplyr::select(AgricultureAgriculture:length(.)) %>% tbl_df
  discEntropy <- data_frame(discBreadth = rowSums(-(b * log(b)),na.rm = TRUE), unitID = academicsBF$unitID) %>% mutate(BF_discBreadth = discBreadth - median(discBreadth))
  academicsBF %<>% inner_join(discEntropy,by='unitID')
  
  
  # NOW: Build the model covariate data frame with all possible covariate
  # candidates. Each student's model only uses a subset of these columns.  See
  # function 'utility'....
  studentBF <- student     %>% dplyr::select(unitID,College,UGDS,ADM_RATE) %>%
    inner_join(incomeBF    %>% dplyr::select(-College,-UGDS), by='unitID') %>%
    inner_join(ethnicBF    %>% dplyr::select(-College),       by='unitID') %>%
    inner_join(aidBF       %>% dplyr::select(-College,-UGDS), by='unitID') %>%
    inner_join(settingBF   %>% dplyr::select(-College,-UGDS,-probSchool), by='unitID') %>%
    inner_join(earnRepayBF %>% dplyr::select(-College,-UGDS,-probSchool), by='unitID') %>%
    #  inner_join(costBF      %>% dplyr::select(-College,-UGDS), by='unitID') %>%
    inner_join(academicsBF %>% dplyr::select(-College,-UGDS), by='unitID') %>%
    mutate(BF_prior = log10(nrow(.)*probSchool)) 
  
  #     }
  #   }
  cat("Done.\n")
  
  # This maps the student profile property names to the college Bayes factor
  # data table ('studentBF') column names.
  propertyMap <- list(
    ethnicity  = c(white    = 'BF_WHITE',
                   black    = 'BF_BLACK',
                   hispanic = 'BF_HISP',
                   asian    = 'BF_ASIAN'),
    income     = list(dependent   = c(le30K       = "BF_le30K",
                                      gt30Kle110K = "BF_gt30Kle110K",
                                      gt110K      = "BF_gt110K"),
                      independent = c(le30K       = "BF_le30K",
                                      gt30Kle110K = "BF_gt30Kle110K",
                                      gt110K      = "BF_gt110K")),
    sat        = c(le800        = 'BF_SAT_le800',  
                   gt800le1000  = 'BF_SAT_gt800le1000', 
                   gt1000le1200 = 'BF_SAT_gt1000le1200', 
                   gt1200le1400 = 'BF_SAT_gt1200le1400', 
                   gt1400='BF_SAT_gt1400'),
    discipline = setNames(paste0('BF_',discNames),discNames),
    discBreadth= c(entropy = 'BF_discBreadth'),
    fafsa      = c(fsend_1 = 'BF_fsend_1_2005',
                   fsend_2 = 'BF_fsend_2_2005',
                   fsend_3 = 'BF_fsend_3_2005',
                   fsend_4 = 'BF_fsend_4_2005',
                   fsend_5 = 'BF_fsend_5_2005'),
    region     = c(FarWest        = 'BF_FarWest(AK,CA,HI,NV,OR,WA)',
                   GreatLakes     = 'BF_GreatLakes(IL,IN,MI,OH,WI)',
                   MidEast        = 'BF_MidEast(DE,DC,MD,NJ,NY,PA)',
                   NewEngland     = 'BF_NewEngland(CT,ME,MA,NH,RI,VT)',
                   Plains         = 'BF_Plains(IA,KS,MN,MO,NE,ND,SD)',
                   RockyMountains = 'BF_RockyMountains(CO,ID,MT,UT,WY)',
                   Southeast      = 'BF_Southeast(AL,AR,FL,GA,KY,LA,MS,NC,SC,TN,VA,WV)',
                   Southwest      = 'BF_Southwest(AZ,NM,OK,TX)'),
    locale     = c(Rural          = 'BF_localeAggRural',
                   TownRemote     = 'BF_localeAggTownRemote',
                   TownDistant    = 'BF_localeAggTownDistant',
                   SuburbSmallMid = 'BF_localeAggSuburbSmall/Midsize & Town:Fringe',
                   SuburbLarge    = 'BF_localeAggSuburbLarge',
                   CitySmall      = 'BF_localeAggCitySmall',
                   CityMidsize    = 'BF_localeAggCityMidsize',
                   CityLarge      = 'BF_localeAggCityLarge'),
    earnings   = c(le30K       = 'BF_p_le30K',  
                   gt30Kle48K  = 'BF_p_gt30Kle48K', 
                   gt48Kle75K  = 'BF_p_gt48Kle75K', 
                   gt75Kle110K = 'BF_p_gt75Kle110K', 
                   gt110K      = 'BF_p_gt110K' ),
    dfltrate   = c(cdr3yr      = 'BF_CDR3est'),
    prior      = 'BF_prior',
    completion = 'BF_C150_4_POOLED_SUPP',
    admission  = 'BF_ADM_RATE',
    gender     = c(female = 'BF_female_2005',
                   male   = 'BF_female_2005')
  )
  
  makeTraitLabels <- function(trait,ntrait,levels){
    # Note that this must be called with positive integers for 'trait' &
    # 'ntrait' and ntrait <= length(levels) (with intention to use
    # ntrait==length(levels)) where 'levels' is a list of character vectors. 
    # Will return a character vector with length = prod(sapply(levels,length)).
    if(trait>=ntrait) return(levels[[trait]])
    return(c(sapply(levels[[trait]],paste0,makeTraitLabels(trait+1,ntrait,levels))))
  }
  traitLevels  <- list(
    Risk      = c('Lr','Mr','Hr'),
    Vision    = c('Lv','Mv','Hv'),
    Breadth   = c('Lb','Mb','Hb'),
    Challenge = c('Lc','Mc','Hc'))
  traitLabels  <- makeTraitLabels(trait=1,ntrait=length(traitLevels),levels=traitLevels)
  StudentTrait <- setNames(strsplit(traitLabels,'[a-z]'),traitLabels)
  
  #   show(studentBF %>% 
  #        filter(grepl('Harvard|Northwestern U|Mass.+Inst.+Tech.+|Cal.+Inst.+Tech.+',College)) %>% 
  #        dplyr::select(1:5,BF_Engineering,BF_localeAggCityLarge) %>% 
  #        as.data.frame)
  return(list(studentBF   = studentBF,
              student     = student,
              propertyMap = propertyMap,
              StudentTrait= StudentTrait,
              discgrp     = discgrp
  )
  )
}
```


```{r functions,echo=FALSE}
#================================================================
# Function getParameters:
getParameters <- function(studentProfile,propertyMap){
  # Get the coefficients for the student profile:
  propCategory <- c('SelfOrigin','Setting','Academics','Prospects')
  traitLvls    <- setNames(as.list(studentProfile$traits),c('Risk','Vision','Breadth','Challenge'))
  propsByTrait <- list(Risk      = list(SelfOrigin = c('ethnicity','income'),
                                        Setting    = c('region','locale')), 
                       Vision    = list(Prospects  = c('completion','earnings','dfltrate'),
                                        Setting    = c('region','locale')), 
                       Breadth   = list(Academics  = c('sat','discipline','discBreadth')), 
                       Challenge = list(Academics  = c('sat','admission','discipline'),
                                        SelfOrigin = c('ethnicity','income')))
  
  # Initialize beta with weightings of 1.0.
  # beta <- matrix(rep(1,length(xBF)),ncol=1,dimnames=list(c('ethnicity','income','sat','fafsa'),'Utility'))
  beta <- matrix(rep(1,length(propertyMap)),ncol=1)
  rownames(beta) <- names(propertyMap)
  
  # Initial weightings relative to a full weighting value of 1.0:
  beta['prior'    ,1] <-  0.5  # downweight size of school
  beta['admission',1] <- -0.5  # prefer schools that have lower admissions rates (i.e., more selective)
  beta['income'   ,1] <-  0.5  # downweight income
  beta['locale'   ,1] <-  0.8  # downweight locale
  beta['fafsa'    ,1] <-  0.3  # downweight fafsa
  beta['dfltrate' ,1] <- -0.5  # prefer schools with lower 3-yr credit default rates (CDR3)
  # assumes stronger influence of gender for women than men...
  beta['gender'   ,1] <- ifelse(studentProfile$gender == 'female',0.5,-0.1) 
  
  # (^_^) ...there's gotta be a better way of doing this...
  for(pCat in propCategory){
    prpnm <- propsByTrait$Risk[[pCat]]
    if (!is.null(prpnm)){
      # must have vectors for each trait level with as many elements as are in the vectors of the propsByTrait list
      beta[prpnm,1] <- beta[prpnm,1] * switch(traitLvls$Risk,
                                              L=list(SelfOrigin = c(ethnicity=3.0,income=2.0),
                                                     Setting    = c(region   =3.0,locale=2.0))[[pCat]],
                                              M=list(SelfOrigin = c(ethnicity=0.3,income=0.3),
                                                     Setting    = c(region   =1.0,locale=1.0))[[pCat]],
                                              H=list(SelfOrigin = c(ethnicity=0.1,income=0.1),
                                                     Setting    = c(region   =0.1,locale=0.1))[[pCat]])
    }
    prpnm <- propsByTrait$Vision[[pCat]]
    if (!is.null(prpnm)){
      beta[prpnm,1] <- beta[prpnm,1] * switch(traitLvls$Vision,
                                              L=list(Prospects = c(completion=0.1,earnings=0.1, dfltrate=0.2),
                                                     Setting   = c(region    =1.5,locale  =1.5))[[pCat]],
                                              M=list(Prospects = c(completion=1.0,earnings=0.5, dfltrate=1.0),
                                                     Setting   = c(region    =1.0,locale  =1.0))[[pCat]],
                                              H=list(Prospects = c(completion=3.0,earnings=2.0, dfltrate=3.0),
                                                     Setting   = c(region    =0.5,locale  =0.5))[[pCat]])
    }
    prpnm <- propsByTrait$Breadth[[pCat]]
    if (!is.null(prpnm)){
      beta[prpnm,1] <- beta[prpnm,1] * switch(traitLvls$Breadth,
                                              L=list(Academics=c(sat=1.5,discipline=3.5,discBreadth= -0.3))[[pCat]],
                                              M=list(Academics=c(sat=1.0,discipline=1.0,discBreadth=  0.3))[[pCat]],
                                              H=list(Academics=c(sat=1.5,discipline=0.3,discBreadth=  1.0))[[pCat]])
    }
    prpnm <- propsByTrait$Challenge[[pCat]]
    if (!is.null(prpnm)){
      beta[prpnm,1] <- beta[prpnm,1] * switch(traitLvls$Challenge,
                                              L=list(Academics  = c(sat=3.0,admission=0.1,discipline=2.0),
                                                     SelfOrigin = c(ethnicity=2.0,income=2.0))[[pCat]],
                                              M=list(Academics  = c(sat=1.0,admission=1.0,discipline=1.0),
                                                     SelfOrigin = c(ethnicity=1.3,income=1.3))[[pCat]],
                                              H=list(Academics  = c(sat=3.0,admission=3.0,discipline=1.3),
                                                     SelfOrigin = c(ethnicity=1.0,income=1.0))[[pCat]])
    }
  }
  
  # Also, make the beta's of traits sum to 1 at each level so that final utility
  # is roughly same magnitude as a typical Bayes factor for school properties.
  # This facilitates interpretation of utility in similar fashion as a raw Bayes factor.
  beta[,1] <- beta[,1]/sum(abs(beta[,1]))
  
  return(beta)
}
#================================================================
# Function plotBeta to display the student's utility partworths.
plotBeta <- function(beta,stdtProf){
  trtstr <- paste(paste(c('Risk','Vision','Breadth','Challenge'),
                        stdtProf$traits,sep=":"),collapse=', ')
  b <- data_frame(Trait=factor(rownames(beta),
                               levels=rownames(beta)[order(abs(beta[,1]),decreasing=TRUE)]),
                  Value=beta[,1])
  gplt <- b %>% 
    ggplot(aes(x=Trait,y=Value,fill=Trait)) 
  gplt <- (gplt + 
             geom_bar(stat='identity',position='dodge') + 
             #theme(text=element_text(face='bold')) + 
             scale_y_continuous(lim=c(-1,1)*3/4) +
             ggtitle(sprintf("Utility Partworths for %s",trtstr))) %T>% print
  invisible(gplt)
}
#================================================================
# Function plotContributions to display partworth-weighted contributions of the factors for the top schools.
plotContributions <- function(caseResult,plot.it=TRUE){
  # Show the partworth-weighted-covariate contributions to the utility function for each college:
  # Sort the factors into largest to smallest wrt contribution to top school.
  
  iorder <- order(as.data.frame(caseResult$topN)[1,-(1:4)],decreasing = TRUE)
  nms    <- names(caseResult$topN)[-(1:4)][iorder]
  gplt <- caseResult$topN %>% 
    dplyr::select(-c(1:2,4)) %>%
    gather(key=Factor,value=Utility,-labels) %>% 
    mutate(Factor = factor(Factor,levels=nms)) %>%
    ggplot(aes(x=Factor,y=Utility,fill=Factor))
  gplt <- gplt + 
    geom_bar(stat='identity') +
    facet_wrap(~labels) + 
    ggtitle("Utility Contributions of Factors") +
    theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0),
          legend.position = 'none')
  if(plot.it) print(gplt)
  invisible(gplt)
}
#===================================================================
# Function utility:
# Here's the Decision Analysis Model; Probabilistic, Utility-Based Discrete Choice Model
utility <- function(sBF,studentProfile,propertyMap,dump.covariates=FALSE){
  # studentBF must come in as a 1-row matrix!!!
  
  covarnames <- sapply(names(propertyMap),function(propnm){
    if(propnm == 'income'){
      return(propertyMap$income[[ifelse(studentProfile$dependent,
                                        'dependent',
                                        'independent')]][studentProfile[[propnm]]])
    }
    levelName <- studentProfile[[propnm]]
    if(is.null(levelName)) return(propertyMap[[propnm]]) 
    return(propertyMap[[propnm]][levelName])
  })
  if(is.null(dim(sBF))) {
    xBF <- matrix(sBF[covarnames],nrow=1,dimnames=list(NULL,covarnames))
  } else {
    xBF <- matrix(sBF[,covarnames],nrow=1,dimnames=list(NULL,covarnames))
  }
  
  # Assign useful column names...
  colnames(xBF) <- paste(colnames(xBF),names(covarnames),sep=':')
  
  #show(xBF)
  
  # Compute the utility the student derives from the school.
  #   show(studentProfile$beta)
  #   show(colnames(xBF))
  beta <- setNames(studentProfile$beta,colnames(xBF))
  #   beta <- c(ethnicity = 0.008976661, income = 0.008976661, sat = 0.403949731, discipline = 0.035008977, 
  #             fafsa = 0.026929982, region = 0.008976661, locale = 0.004488330, earnings = 0.044883303, 
  #             prior = 0.089766607, completion = 0.089766607, admission = -0.269299820, gender = -0.008976661)
  #show(beta)
  u    <- xBF %*% beta
  
  if (dump.covariates){
    wxBF <- xBF * matrix(beta,nrow=nrow(xBF),ncol=ncol(xBF),byrow = TRUE)
    colnames(wxBF) <- colnames(xBF)
    return(wxBF)
  } else {
    return(u[1])
  }
}

#================================================================
# Function studentCaseStudy to assess the suitability of colleges for a specific student's profile:
studentCaseStudy <- function(caseSpec,dspec,ntop=20,signifLevel=0.15,verbose=FALSE,exact=FALSE){
  
  stdtProf  <- caseSpec$studentProfile
  utility   <- dspec$utility
  
  utilities <- dspec$studentBF %>%
    dplyr::select(starts_with('BF')) %>%
    apply(1,utility,studentProfile=stdtProf,propertyMap=dspec$propertyMap) %>%
    setNames(dspec$studentBF$College )
  #show(sort(utilities,decreasing = TRUE)[1:ntop])
  
  tmp <- dspec$studentBF %>% 
    dplyr::select(starts_with('BF')) %>%
    slice(1) %>% as.matrix %>%
    utility(studentProfile=stdtProf,propertyMap=dspec$propertyMap,dump.covariates=TRUE)
  
  wxBF <- dspec$studentBF %>% 
    dplyr::select(starts_with('BF')) %>%
    apply(1,utility,studentProfile=stdtProf,propertyMap=dspec$propertyMap,dump.covariates=TRUE) %>%
    t %>% as.data.frame %>% tbl_df %>%
    setNames(gsub('^[^:]+:','',colnames(tmp))) %>%
    mutate(unitID  = dspec$studentBF$unitID,
           College = dspec$studentBF$College,
           Utility = utilities) %>% 
    arrange(desc(Utility)) %>%
    dplyr::select(unitID,College,Utility,everything())
  
  #   stdtProf$beta %>% print
  #   wxBF %>% print(width=Inf)
  
  labels <- sprintf("%d. %s",order(order(utilities,decreasing=TRUE)),names(utilities))
  
  stdtUtility <- data_frame(unitID  = dspec$studentBF$unitID, 
                            College = names(utilities),
                            Utility = utilities) %>%
    mutate(labels = factor(labels,levels=labels[order(Utility,decreasing=TRUE)]),
           expChoice = 10^Utility,
           ChoicePct = 100*expChoice/sum(expChoice,na.rm=TRUE)) %>%
    dplyr::select(-expChoice) %>%
    inner_join(wxBF %>% dplyr::select(-College,-Utility),by='unitID') %>% 
    arrange(desc(Utility)) #%T>% print
  
  topN <- stdtUtility %>% dplyr::select(-College) %>% top_n(ntop,Utility) 
  
  if(verbose) {
    topN %>% print
    traitStr <- paste(paste(c('Risk','Vision','Breadth','Challenge'),stdtProf$traits,sep=":"),collapse=', ')
    profText <- stdtProf[!grepl('beta|traits',names(stdtProf))]
    pltTitle <- paste(strwrap(paste(paste(names(profText),profText,sep=':'),collapse=', '),width = 80),collapse='\n')
    pltTitle <- paste0("Top ",ntop," Colleges for Profile:\n",pltTitle,'\n',traitStr)
    
    gplt0 <- topN %>% 
      mutate(labels = factor(labels,levels=rev(levels(labels))),
             labelY = pmin(0,min(Utility))) %>%
      ggplot(aes(x=labels,y=Utility,fill=labels))
    gplt <-  gplt0 +
      geom_bar(stat='identity',position='dodge') + coord_flip() + 
      geom_text(aes(x=labels,y=labelY,label=labels),size=6,hjust=0,vjust=0.5) +
      theme(text=element_text(face='bold',size=8),
            axis.text.y=element_blank(),
            title=element_text(size=12,face='bold'),
            legend.position = "none") +
      ggtitle(pltTitle) +
      labs(x=sprintf("Top %d Colleges",ntop)) 
    gplt %>% print
  }
  
  # Back-calculate the beta coefficients and see if can determine which
  # covariate contributed most strongly to the utility rankings.
  stdtBF <- stdtUtility %>% filter(complete.cases(.))  %>%
    inner_join(dspec$studentBF %>% dplyr::select(-College),by='unitID')
  axbf <- stdtBF %>% dplyr::select(one_of(gsub('^[^:]+:','',colnames(tmp)))) %>% as.matrix
  axbf <- axbf[, apply(axbf,2,function(x) sd(x,na.rm = TRUE)>0) ]
  betacoeff <- solve(t(axbf) %*% axbf,t(axbf) %*% matrix(stdtBF$Utility,ncol=1))
  
  if(verbose) show(betacoeff[betacoeff>1.0E-9,])
  
  stdtBF %<>%
    dplyr::select(c(4,3,5),one_of(names(betacoeff[abs(betacoeff)>1.0E-9,]))) 
  
  if(verbose) stdtBF %>% print(n=ntop,width=Inf)
  
  utlBFcor <- stdtBF %$% 
    lapply(.[-(1:3)],function(y) cor.test(Utility[1:ntop],y[1:ntop],method='spearman',exact=exact)) #%T>% print
  
  # This shows which covariates dictate the rankings by utility. If significant
  # negative correlation, then that covariate is important in the latter half of
  # the ranking (note that reverse ordering due to negative coefficients is
  # already accounted for).
  pvals <- sapply(utlBFcor,'[[','p.value')
  pvals <- pvals[!is.na(pvals)]
  
  signifCor <- utlBFcor[pvals <= pmax(signifLevel,min(pvals,na.rm=TRUE))] %$% {.[order(sapply(.,'[[','p.value'))]}
  
  if(verbose) signifCor %>% show
  
  results <- list(studentProfile=stdtProf,ntop=ntop,topN=topN,utilities=stdtUtility,BF=stdtBF,cor=utlBFcor,signifCor=signifCor,beta=betacoeff)
  invisible(results)
}
#================================================================
plotTopN <- function(caseResult,keyprop='region',plot.it=TRUE){
  propnm   <- keyprop
  stdtProf <- caseResult$studentProfile
  ntop     <- nrow(caseResult$topN)
  profText <- c(stdtProf[propnm],stdtProf[!grepl(paste0('beta|traits|',propnm),names(stdtProf))])
  pltTitle <- paste(strwrap(paste(paste(names(profText),profText,sep=':'),collapse=', '),width = 80),collapse='\n')
  traitStr <- paste(paste(c('Risk','Vision','Breadth','Challenge'),stdtProf$traits,sep=":"),collapse=', ')
  pltTitle <- paste0("Top ",ntop," Colleges for Profile:\n",pltTitle,'\n',traitStr)
  
  gplt0 <- caseResult$topN %>% 
    mutate(labels = factor(labels,levels=rev(levels(labels))),
           labelY = pmin(0,min(Utility))) %>%
    ggplot(aes(x=labels,y=Utility,fill=labels))
  gplt <-  gplt0 +
    geom_bar(stat='identity',position='dodge') + coord_flip() + 
    geom_text(aes(x=labels,y=labelY,label=labels),size=6,hjust=0,vjust=0.5) +
    theme(text=element_text(face='bold',size=8),
          axis.text.y=element_blank(),
          title=element_text(size=12,face='bold'),
          legend.position = "none") +
    ggtitle(pltTitle) +
    labs(x=paste("Top",ntop,"Colleges"),sep=" ")
  
  if(plot.it) print(gplt)
  invisible(gplt)
  
}
#================================================================
# Function sensitivity to do sensitivity analysis (parametric sweeps)
sensitivity <- function(stdtProf0,pmap,caseResult,spec,maxcnt=Inf,plot.it=TRUE){
  studentBF    <- spec$studentBF
  StudentTrait <- spec$StudentTrait
  propertyMap  <- spec$propertyMap
  
  propnms <- names(pmap)
  cnt <- 0
  
  caseList <- list()
  grobList <- list()
  for(propnm in propnms){
    stdtProf <- stdtProf0
    # par(mfrow=c(2,ceiling(length(pmap[[propnm]])/2)))
    if(propnm == 'income') {
      if(stdtProf$dependent) {
        lvls <- names(pmap[[propnm]]$dependent)
      } else {
        lvls <- names(pmap[[propnm]]$independent)
      }
    } else {
      lvls <- names(pmap[[propnm]])
    }
    for(levelnm in lvls){
      stdtProf[[propnm]] <- levelnm
      stdtProf$beta <- getParameters(stdtProf,spec$propertyMap)
      caseResult$studentProfile <- stdtProf
      caseResult    <- studentCaseStudy(caseResult,spec,ntop = 10,verbose = FALSE)
      
      profText <- c(stdtProf[propnm],stdtProf[!grepl(paste0('beta|traits|',propnm),names(stdtProf))])
      pltTitle <- paste(strwrap(paste(paste(names(profText),profText,sep=':'),collapse=', '),width = 80),collapse='\n')
      traitStr <- paste(paste(c('Risk','Vision','Breadth','Challenge'),stdtProf$traits,sep=":"),collapse=', ')
      pltTitle <- paste0("Top 10 Colleges for Profile:\n",pltTitle,'\n',traitStr)
      
      nms <- names(caseList)
      caseList <- setNames(c(caseList,list(caseResult)), c(nms,pltTitle))
      
      #if(plot.it){
      gplt0 <- caseResult$topN %>% 
        mutate(labels = factor(labels,levels=rev(levels(labels))),
               labelY = pmin(0,min(Utility))) %>%
        ggplot(aes(x=labels,y=Utility,fill=labels))
      gplt <-  gplt0 +
        geom_bar(stat='identity',position='dodge') + coord_flip() + 
        geom_text(aes(x=labels,y=labelY,label=labels),size=6,hjust=0,vjust=0.5) +
        theme(text=element_text(face='bold',size=8),
              axis.text.y=element_blank(),
              title=element_text(size=12,face='bold'),
              legend.position = "none") +
        ggtitle(pltTitle) +
        labs(x="Top 10 Colleges") 
      #gplt %>% print
      grobList <- c(grobList,list(gplt))
      #}
      
      cnt <- cnt + 1
      if(cnt>=maxcnt) break
    }
    if(cnt>=maxcnt) break
  }
  # par(paropts)
  
  ### UNFORTUNATELY, SEEMS TO LOSE CONTROL OF TEXT SIZE WHEN ggplots ARE LAID IN
  ### GRID USING PACKAGE 'gridExtra'.... HELP!!!
  #gmulti <- marrangeGrob(grobList,nrow=2,ncol=4)
  if(plot.it) do.call(grid.arrange,grobList)
  return(list(caseList=caseList,grobList=grobList))
}


#================================================================
# Function plotMap to plot an interactive map of the top schools.
plotMap <- function(caseResult,dspec){
  mortarBoardIcon <- makeIcon(
    iconUrl = "https://upload.wikimedia.org/wikipedia/commons/thumb/7/76/Fxemoji_u1F393.svg/512px-Fxemoji_u1F393.svg.png",
    iconWidth = 30, iconHeight = 30)
  
  schools <- caseResult$topN %>%
    inner_join(dspec$student %>% 
                 dplyr::select(c(unitID,College,city,state,zip,median_hh_inc_2005,lon,lat)), by='unitID') %>%
    mutate(info = paste0(College,'<br>',city,', ',state,' ',zip,
                         sprintf('<br> Rank: %d Utility: %.2f',seq_along(Utility),Utility)))
  map <- leaflet(schools) %>% 
    setView(-93.65, 42.0285, zoom = 4) %>%
    addTiles() %>%
    addMarkers(~lon, ~lat, popup=~info,
               options = popupOptions(closeButton = TRUE),
               clusterOptions = markerClusterOptions(), 
               icon = mortarBoardIcon)
  #print(map)
  map
}

```


```{r makedataspec, warning=FALSE,message=FALSE,results='hide'}
#================================================================
# RUN THE SETUP FUNCTION:
DataSpec <- make_DataSpec()
DataSpec$utility <- utility

#================================================================

# # Assign setup variables to objects in current environment:
# for( vnm in names(setupResult )){
#   eval(parse(text=sprintf('%s = setupResult$%s',vnm,vnm)))
# }

```

# Here are ***Your*** Top Colleges

Apply the model to get your `ntop` colleges: 

```{r topn,warning=FALSE,tidy=TRUE}
# ENTER N, for top-N colleges:
Case$ntop <- 20
Case$studentProfile$beta <- getParameters(Case$studentProfile,DataSpec$propertyMap)
Case <- studentCaseStudy(Case,DataSpec,verbose=FALSE,ntop=Case$ntop)

# # Check-in on some select schools:
# selectSchools <- paste(c('Massachusetts Institute of Technology', 'California Institute of Technology',
#                          'Princeton', 'Yale', 'Harvard', 'Stanford', 'Duke', 'Vanderbilt', 'Berkeley',
#                          'Northwestern U', 'Princeton', 'Cornell U'),collapse="|")
# Case$BF %>% 
#   filter(grepl(selectSchools,labels)) %$% {
#     tmp<-t(.[-1]); 
#     dimnames(tmp)<-list(names(.)[-1],gsub(' | University|nstitute of |echnology|ifornia|assachusetts|of','',labels)); 
#     tmp
#     } %>% 
#   round(2)
```

Display the results:

```{r topNplot,fig.width=10,fig.height=8}
plotTopN(Case,plot.it = TRUE)
```

Display a bit more output to help explain why those are the top schools. 

The factor `prior` refers to the contribution due to size of school in terms of undergraduate enrollment (database field `"UGDS"`).

```{r topNexpln,warning=FALSE,fig.width=10,fig.height=8}
# Plot the partworth contributions to the utility for each of the top schools.
plotContributions(Case,plot.it = TRUE)

# Show the partworth (utility function parameters):
plotBeta(Case$studentProfile$beta,Case$studentProfile)

```


# How About a Road-Trip to Your Top Schools?!
I've copied idea and code snips from Tad Dallas's script on Kaggle.com:

```{r roadtrip, eval=TRUE, echo=FALSE, tidy=TRUE}
# Display an interactive map highlighting the top schools.
plotMap(Case,DataSpec)
```

# Sensitivity of Your Top Schools to Changes in Your Profile
Perform sensitivity analysis on your profile factors.

```{r snstvty, warning=FALSE,fig.width=10,fig.height=8,tidy=TRUE}
# Sweep region
sweepVariable <- 'region'
case0       <- Case
stdtProf0   <- case0$studentProfile
pmap        <- DataSpec$propertyMap[sweepVariable]
maxcnt      <- length(pmap[[sweepVariable]])
sensResults <- sensitivity(stdtProf0,pmap,case0,DataSpec,maxcnt=maxcnt,plot.it=FALSE)
print(sensResults$grobList)

# This is how to sweep over SAT:
sweepVariable <- 'sat'
case0       <- Case
stdtProf0   <- case0$studentProfile
pmap        <- DataSpec$propertyMap[sweepVariable]
maxcnt      <- length(pmap[[sweepVariable]])
sensResults <- sensitivity(stdtProf0,pmap,case0,DataSpec,maxcnt=maxcnt,plot.it=FALSE)
print(sensResults$grobList)
# 
# sweepVariable <- 'locale'
# stdtProf0   <- studentProfile
# pmap        <- propertyMap[sweepVariable]
# maxcnt      <- length(pmap[[sweepVariable]])
# sensResults <- sensitivity(stdtProf0,pmap,Case,DataSpec,maxcnt=maxcnt,plot.it=FALSE)
# print(sensResults$grobList)

# # This is how to sweep over a subset of disciplines:
# stdtProf0 <- Case$studentProfile
# pmap      <- DataSpec$propertyMap[c('discipline')]
# pmap$discipline <- pmap$discipline[c('ComputerInformation','BiologicalBiomedical','HealthProfessions','PhysicalSciences','MathematicsStatistics')]
# maxcnt          <- pmin(7,length(pmap$discipline)) 
# sensResults     <- sensitivity(stdtProf0,pmap,Case,DataSpec,maxcnt=maxcnt,plot.it=FALSE)
# print(sensResults$grobList)

# # This is how to sweep over a behavioral trait:
# stdtProf0 <- Case$studentProfile
# ntop <- 10
# for(lvl in c("L","M","H")){
#   stdtProf0$traits$Risk <- lvl
#   stdtProf0$beta <- getParameters(stdtProf0,DataSpec$propertyMap)
#   case0          <- list(studentProfile=stdtProf0,ntop=ntop)
#   case0          <- studentCaseStudy(case0,DataSpec,verbose=FALSE,ntop=ntop)
#   plotTopN(caseResult = case0,plot.it = TRUE)
# }

```

# Explanation of Valid Entries for Student Profile
The following selections show what entries are valid for the student's `region`, `locale`, `discipline` and behavioral traits.

## Valid Regions
Valid region names to choose for your student profile `region` property are shown in column ***Region*** of the table below; and states in each region are listed in the ***States*** column:

```{r region, echo=FALSE,results='asis'}
regions <- data.frame(Region = sort(names(DataSpec$propertyMap$region)),
                      States = gsub('[^(]+\\(([^)]+)).*','\\1',
                                    DataSpec$propertyMap$region[sort(names(DataSpec$propertyMap$region))]))
rownames(regions) <- seq_len(nrow(regions))
print(xtable(regions),type='html')

```

## Valid Locales
Valid locale names to choose for your student profile `locale` property are shown in column ***Locale*** of the table below:

```{r locale, echo=FALSE,results='asis'}
#student %$% { locale %>% summary %>% print }

localeNames <- sort(unique(as.character(DataSpec$student$locale)))
descript <- setNames(c(gsub('([^ ]+) ([^ ]+).+','\\1\\2',localeNames[1:3]),
                       rep('Rural',3),
                       gsub('([^ ]+) ([^ ]+).+','\\1\\2',localeNames[7]),
                       rep('SuburbSmallMid',2),
                       gsub('([^ ]+) ([^ ]+).+','\\1\\2',localeNames[10]),
                       'SuburbSmallMid',
                       gsub('([^ ]+) ([^ ]+).+','\\1\\2',localeNames[12])),localeNames)
descript <- gsub('[[:punct:]]','',descript)

locales <- data.frame(Locale = descript,Description = names(descript))
locales <- locales[c(6,4,5,12,10,11,9,8,7,3,2,1),]
rownames(locales) <- seq_len(nrow(locales))
print(xtable(locales),type='html')

```

## Valid Academic Disciplines
Valid academic disciplines to choose for your student profile `discipline` property are shown in column ***discName*** of the table below:

```{r discipline, echo=FALSE,results='asis'}

#print(xtable(DataSpec$discgrp[c("LABEL","discName")]),type='html')
DataSpec$discgrp %>% 
  arrange(discName) %>% 
  dplyr::select(LABEL,discName) %>% 
  xtable %>% 
  print(type='html')
```

## Explanation of Behavioral Traits

Trait level has 4 types with impact on properties as follows when each trait increases from low to medium to high: 

1. Risk      -- Willingness to surround self with people & settings very different from self & origins;
+ Weights homogeneity & sameness as self lower than other traits.
2. Vision    -- Willingness to look past short-term suitability in pursuit of future;
+ Weights completion rates, future earnings, debt & debt repayment rates higher than current campus setting compatability.
3. Breadth   -- Willingness to entertain a variety of academic disciplines;
+ Weights predominance of own major discipline lower than otherwise. 
4. Challenge -- Willingness to embrace highest academic rigor
+ Weights admissions selectiveness, high SAT & higher than otherwise.



Traits Have Impact On Four Property Categories Listed in Rows:

|                              |      Risk       |      Vision       |      Breadth       |      Challenge       |      Properties
|  ----------------------------|-----------------|-------------------|--------------------|----------------------|---------------------------
|   1. Self & Origins          | same to differ  |      ------       |      -------       |        ----          | ethn,inc,1stgen,usbrn,pov
|   2. Campus Setting & Folks  |homog to diverse |      ------       |      -------       |        ----          | reg,loc,siz
|   3. Academics               |      ----       |      ------       |  narrow to broad   | low to high sat/adm  | sat,admrt,disc
|   4. After-College Prospects |      ----       |low to high dbt/ern|      -------       |        ----          | comprt,earn,rpy,dfltrt


Any feedback would be appreciated.

Thanks,

-Michael L. Thompson 

[(my info at LinkedIn)](<https://www.linkedin.com/in/mlthomps>)


