---
title: "Which College is Best for You?"
author: "Michael L. Thompson"
date: "January 6, 2016"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---
# Introduction
The [US College Scorecard dataset](<https://collegescorecard.ed.gov>)
provides everyday folks, like myself, access to a rich source of US college
information.  Yet, it's quite a challenge making sense out of such a
sprawling dataset.  And even moreso to make sense out of it in a way that
captures the specific context of a student's interest in determining which
colleges are good candidates to attend or not.

Before jumping into the analysis, I'd like to acknowledge those folks who
generously shared their scripts on [Kaggle.com](<https://www.kaggle.com>)
from which I learned -- and outright copied -- code snips to get this done. 
In particular, the code provided by Ben Hamner on using package 'RSQLite' and
by Tad Dallas on using package 'leaflet' for map plotting. Thank you!!!

## Comments:
***ALSO:*** To see the payoff, go straight to the ***Sensitivity Analysis*** section. 

My intent was to render the R script as a PDF-format notebook using 
RStudio's File-->Compile Notebook... menu command. But it was just taking too long 
for me to get all the formatting of plots, lists and tables right -- so I bailed!

The R script version runs fine when sourced at the R console or in
RStudio. In fact, that's the only way I could get the `leaflet` package's
interactive mapping to display -- until I changed `print(map)` to just plain `map`.

Any help in cleaning up the formatting would be much appreciated.

# Objective
I set out to build a decision analysis tool that would be useful to an
individual, like me, for judging the suitability of colleges specifically
attuned to my criteria.  The analysis is based upon a model that combines the
US College Scorecard data and my personal insights -- i.e., domain expertise
-- into something useful for the following:

1.  Estimating what specific type student is attracted to each college; 
2.  Estimating what specific types of student are most likely to thrive at each college; and 
3.  Predicting whether a student with a specific profile of interests & character traits would thrive at a specific college or not.

This is still but a preliminary proof-of-concept level of analysis to
illustrate the possibilities with this framework. Ideally, folks will see
this, revise, complete and extend it.

# Modeling Plan
My modeling plan looks like this:

1. Explore the dataset with unsupervised learning and data visualization to understand some of the underlying structure of the data.
+    a) Define contexts describing (i) College Experience, (ii) Post-College Professional Success, (iii) Post-College Life Satisfaction in terms of sets of variables in the Scorecard dataset related to salient aspects of each context (e.g., College Experience = f(student origins, student demographics, student performance, college environ/community/campus experience, college academic rigor); and noting that the different contexts may share subsets of variables (e.g. Post-College Life Satisfaction = f(degree achievement, post-college success, earnings vs. debt burden & debt repayment)?
+    b) Predict what colleges cluster together in what Contexts?
2. Create archetypical behavioral characteristics of students and map those behavioral traits to variables in the Scorecard dataset.
3. Build probabilistic choice models that estimate which colleges maximizes a specific student's utility according to the
students personalize valuation coefficients (i.e. partworths).
4. Validate (with personal experience/anecdotal evidence) whether or not specific profiles predicted to thrive at specific colleges in each context make sense or not.

# Data Pre-Processing & Exploration
I've done a ton of pre-processing, plotting and data transformations, even
unsupervised learning, to explore the data.  I'm going to skip the
presentation of those exploration pieces here but instead just show the data
winnowing and wrangling I did in preparation for modeling.

Let's start by loading the packages, downloading & installing them if necessary:
```{r}
library('RSQLite')   # for SQLite manipulation of database
library('magrittr')  # for piping with infix %>%, %<>%, %T>% (for cool, sophistication effect)
library('tidyr')     # for reshaping data (e.g., 'gather')
library('dplyr')     # for data wrangling (e.g., 'filter', 'select', 'summarize', 'inner_join')
library('ggplot2')   # for best-in-class visualization
library('gridExtra') # for grid layout of ggplots
#  library('poLCA')     # for estimation of latent classes underlying categorical variables
library('bnlearn')   # for learning BBN from datasets, and also provides function "discretize"
#  library('gRain')     # for Bayesian belief network (BBN) inference engine
library('leaflet')   # for mapping of the top schools for each student
library('htmltools') # for assistance with leaflet
library('RColorBrewer') # for color palette management
```

After orienting myself a little with the [data documentation](https://collegescorecard.ed.gov/data/documentation/),..., I
started by borrowing code from [Ben Hamner's "Exploring the US College Scorecard Data" script](https://www.kaggle.com/benhamner/us-dept-of-education-college-scorecard/exploring-the-us-college-scorecard-data):

```{r}
db <- dbConnect(dbDriver("SQLite"), "../input/database.sqlite")
# This stops SQLite from writing temp files to disk, which use all the available space
dbGetQuery(db, "PRAGMA temp_store=2;") 

tables <- dbGetQuery(db, "SELECT Name FROM sqlite_master WHERE type='table'")
colnames(tables) <- c("Name")
tables %<>%
  rowwise() %>%
  mutate(RowCount=dbGetQuery(db, paste0("SELECT COUNT(Id) RowCount FROM ", Name))$RowCount[1])
show(tables)

allFields <- dbListFields(db,tables$Name[1])
show(length(allFields))
```
I also downloaded the data dictionary and relied upon it heavily to
understand the database contents. Here I read it in and "revise" it by
getting rid of the strange characters that were meant to be apostrophe's and,
more importantly, filled in the blank cells of its tabular format so it shows
up like a standard data.frame useful for automated processing:
```{r}
if(!exists("dataDict")) {
  if(!file.exists("CollegeScorecardDataDictionary-09-12-2015_REVISED.csv")) {
    # JUST ASSUMES THAT FILE "CollegeScorecardDataDictionary-09-12-2015.csv"
    # DOES EXIST IN THE WORKING DIRECTORY:
    dataDict  <- read.csv(file="../input/CollegeScorecardDataDictionary-09-12-2015.csv",
                          stringsAsFactors = FALSE)
    show(head(dataDict))
    str(dataDict)
    
    show(setdiff(allFields,dataDict$VARIABLE.NAME))
    show(setdiff(dataDict$VARIABLE.NAME,allFields))
    
    dataDict <- dataDict %>% mutate(LABEL=gsub("^(.*)’(.*)$","\\1'\\2",LABEL))
    # Fill-down
    fillDown <- function(df=dataDict,fillCols=c(1:5,8)) {
      nr   <- nrow(dataDict)
      irow <- 1
      while(irow < nr){
        while(df$NAME.OF.DATA.ELEMENT[irow] == ""){
          df[irow,fillCols] <- df[irow-1,fillCols]
          irow <- irow + 1
        }
        irow <- irow + 1
      }
      return(df)
    }
    dataDict <- dataDict %>% fillDown(c(1:5,8))
    #    write.csv(dataDict,file="CollegeScorecardDataDictionary-09-08-2015_REVISED.csv",
    #              row.names = FALSE,na="")
  } else {
    dataDict <- read.csv(file="../input/CollegeScorecardDataDictionary-09-12-2015_REVISED.csv",
                         stringsAsFactors = FALSE)
  }
}
print(head(dataDict))
print(dataDict %>% tbl_df)
```
# Probabilistic Discrete Choice Modeling
A key concept in the probabilistic Bayesian modeling framework I've chosen is to represent each college by a set of variables quantifying the types of students who populate their campus. This requires turning as many of the provided variables as possible into conditional probability distributions such as P(student-income > x | college = school-of-interest). From these conditional probabilities, 

I then computed [Bayes Factors](http://bayesfactor.blogspot.com/2014/02/the-bayesfactor-package-this-blog-is.html): log10(P(student-income > x | college = school-of-interest)/P(student-income > x | college != school-of-interest)) = Posterior-Odds(college = school-of-interest | student-income>x)/Prior-Odds(college = school-of-interest).

These Bayes factors then serve as the covariates in the linear utility function of the discrete choice model: u(school-of-interest | student) = beta(student) * Bayes-factors(school-of-interest).  (See Nobel Laureate for Economics [Daniel McFadden's](https://en.wikipedia.org/wiki/Daniel_McFadden) contributions in discrete choice modeling.)

Then the student can just rank-order the colleges by their utilities u(school-of-interest | student). 

Now, there are many alternatives to this framework.  I chose this one because of its flexibility and my experience using such models. And even though I don't do a rigorous data-estimated hierarchical model using tools such as Stan through package rstan, the promise of the framework is hinted at in the flexibility of the subsequent analyses.

Let's get on with the work!

Grab the requisite datasets: Used years 2005 (for the Treasury data on students' originating zip codes) and 2013.
```{r}
# This groups the disciplines (degrees) into reasonably similar sets.
#discgrp <- read.csv("discgrp.csv", stringsAsFactors=FALSE)
discgrp <- data_frame(LABEL=c("Agriculture, Agriculture Operations, and Related Sciences","Natural Resources and Conservation",
                              "Architecture and Related Services","Area, Ethnic, Cultural, Gender, and Group Studies",
                              "Communication, Journalism, and Related Programs","Communications Technologies/Technicians and Support Services",
                              "Computer and Information Sciences and Support Services","Personal and Culinary Services",
                              "Education","Engineering","Engineering Technologies and Engineering-Related Fields",
                              "Foreign Languages, Literatures, and Linguistics","Family and Consumer Sciences/Human Sciences",
                              "Legal Professions and Studies","English Language and Literature/Letters",
                              "Liberal Arts and Sciences, General Studies and Humanities","Library Science","Biological and Biomedical Sciences",
                              "Mathematics and Statistics","Military Technologies and Applied Sciences","Multi/Interdisciplinary Studies",
                              "Parks, Recreation, Leisure, and Fitness Studies","Philosophy and Religious Studies","Theology and Religious Vocations",
                              "Physical Sciences","Science Technologies/Technicians","Psychology",
                              "Homeland Security, Law Enforcement, Firefighting and Related Protective Services",
                              "Public Administration and Social Service Professions","Social Sciences","Construction Trades",
                              "Mechanic and Repair Technologies/Technicians","Precision Production","Transportation and Materials Moving",
                              "Visual and Performing Arts","Health Professions and Related Programs",
                              "Business, Management, Marketing, and Related Support Services","History"),
                      discrp=c(2,4,2,5,6,4,1,4,5,1,1,5,5,7,5,5,5,2,1,4,5,4,5,5,2,3,3,4,3,3,4,4,4,4,6,2,7,5),
                      VARIABLE.NAME=c("PCIP01","PCIP03","PCIP04","PCIP05","PCIP09","PCIP10","PCIP11","PCIP12","PCIP13",
                                      "PCIP14","PCIP15","PCIP16","PCIP19","PCIP22","PCIP23","PCIP24","PCIP25","PCIP26",
                                      "PCIP27","PCIP29","PCIP30","PCIP31","PCIP38","PCIP39","PCIP40","PCIP41","PCIP42",
                                      "PCIP43","PCIP44","PCIP45","PCIP46","PCIP47","PCIP48","PCIP49","PCIP50","PCIP51",
                                      "PCIP52","PCIP54"))

```

Only deal with colleges having these traits:

1. Public or Private non-profit
   + Drops Private for-profit schools
2. Currently Operating
3. Fully accreditted
4. Predominantly Bachelor's degree granting
5. Within the 50 states of the U.S.A.
6. Not a distance-only college
7. Not a theological school

Now the challenge is to match the available database columns to the criteria I wish to calculate.
```{r}
fieldNames <-
  c(
    'UGDS','UGDS_WHITE','UGDS_BLACK','UGDS_HISP','UGDS_ASIAN','UGDS_AIAN','UGDS_NHPI',
    'UGDS_2MOR','UGDS_NRA','UGDS_UNKN',"UG25abv","INC_PCT_LO","DEP_STAT_PCT_IND","DEP_INC_PCT_LO","IND_INC_PCT_LO",
    "DEP_INC_PCT_M1","DEP_INC_PCT_M2","DEP_INC_PCT_H1","IND_INC_PCT_M1","IND_INC_PCT_M2","IND_INC_PCT_H1","IND_INC_PCT_H2",
    "DEP_INC_PCT_H2","PAR_ED_PCT_1STGEN","pct_white","pct_black","pct_asian","pct_hispanic","pct_ba","pct_grad_prof",
    "pct_born_us","median_hh_inc","poverty_rate","unemp_rate","ln_median_hh_inc","pell_ever","female","fsend_1","fsend_2",
    "fsend_3","fsend_4","fsend_5","INC_PCT_LO","INC_PCT_M1","INC_PCT_M2","INC_PCT_H1","INC_PCT_H2","LATITUDE",
    "LONGITUDE","ZIP","STABBR","region","LOCALE","CCSIZSET","DISTANCEONLY","RELAFFIL","CURROPER","NPT4_PUB",
    "NPT4_PRIV","NPT41_PUB","NPT42_PUB","NPT43_PUB","NPT44_PUB","NPT45_PUB","NPT41_PRIV","NPT42_PRIV","NPT43_PRIV",
    "NPT44_PRIV","NPT45_PRIV","TUITIONFEE_IN","TUITIONFEE_OUT","PCTPELL","ADM_RATE","RET_FT4","SATVR25","SATVR75",
    "SATMT25","SATMT75","SATWR25","SATWR75","SATVRMID","SATMTMID","SATWRMID","ACTCM25","ACTCM75","ACTEN25","ACTEN75",
    "ACTMT25","ACTMT75","ACTWR25","ACTWR75","ACTCMMID","ACTENMID","ACTMTMID","ACTWRMID","SAT_AVG","SAT_AVG_ALL",
    "PCIP01","PCIP03","PCIP04","PCIP05","PCIP09","PCIP10","PCIP11","PCIP12",
    "PCIP13","PCIP14","PCIP15","PCIP16","PCIP19","PCIP22","PCIP23","PCIP24","PCIP25","PCIP26","PCIP27","PCIP29","PCIP30",
    "PCIP31","PCIP38","PCIP39","PCIP40","PCIP41","PCIP42","PCIP43","PCIP44","PCIP45","PCIP46","PCIP47","PCIP48","PCIP49",
    "PCIP50","PCIP51","PCIP52","PCIP54","C150_4_WHITE","C150_4_BLACK","C150_4_HISP","C150_4_ASIAN","ENRL_ORIG_YR2_RT",
    "LO_INC_ENRL_ORIG_YR2_RT","MD_INC_ENRL_ORIG_YR2_RT","HI_INC_ENRL_ORIG_YR2_RT","DEP_ENRL_ORIG_YR2_RT","IND_ENRL_ORIG_YR2_RT",
    "FEMALE_ENRL_ORIG_YR2_RT","MALE_ENRL_ORIG_YR2_RT","PELL_ENRL_ORIG_YR2_RT","NOPELL_ENRL_ORIG_YR2_RT","LOAN_ENRL_ORIG_YR2_RT",
    "NOLOAN_ENRL_ORIG_YR2_RT","FIRSTGEN_ENRL_ORIG_YR2_RT","NOT1STGEN_ENRL_ORIG_YR2_RT","ENRL_ORIG_YR3_RT","LO_INC_ENRL_ORIG_YR3_RT",
    "MD_INC_ENRL_ORIG_YR3_RT","HI_INC_ENRL_ORIG_YR3_RT","DEP_ENRL_ORIG_YR3_RT","IND_ENRL_ORIG_YR3_RT","FEMALE_ENRL_ORIG_YR3_RT",
    "MALE_ENRL_ORIG_YR3_RT","PELL_ENRL_ORIG_YR3_RT","NOPELL_ENRL_ORIG_YR3_RT","LOAN_ENRL_ORIG_YR3_RT","NOLOAN_ENRL_ORIG_YR3_RT",
    "FIRSTGEN_ENRL_ORIG_YR3_RT","NOT1STGEN_ENRL_ORIG_YR3_RT","ENRL_ORIG_YR4_RT","LO_INC_ENRL_ORIG_YR4_RT","MD_INC_ENRL_ORIG_YR4_RT",
    "HI_INC_ENRL_ORIG_YR4_RT","DEP_ENRL_ORIG_YR4_RT","IND_ENRL_ORIG_YR4_RT","FEMALE_ENRL_ORIG_YR4_RT","MALE_ENRL_ORIG_YR4_RT",
    "PELL_ENRL_ORIG_YR4_RT","NOPELL_ENRL_ORIG_YR4_RT","LOAN_ENRL_ORIG_YR4_RT","NOLOAN_ENRL_ORIG_YR4_RT","FIRSTGEN_ENRL_ORIG_YR4_RT",
    "NOT1STGEN_ENRL_ORIG_YR4_RT","ENRL_ORIG_YR6_RT","LO_INC_ENRL_ORIG_YR6_RT","MD_INC_ENRL_ORIG_YR6_RT","HI_INC_ENRL_ORIG_YR6_RT",
    "DEP_ENRL_ORIG_YR6_RT","IND_ENRL_ORIG_YR6_RT","FEMALE_ENRL_ORIG_YR6_RT","MALE_ENRL_ORIG_YR6_RT","PELL_ENRL_ORIG_YR6_RT",
    "NOPELL_ENRL_ORIG_YR6_RT","LOAN_ENRL_ORIG_YR6_RT","NOLOAN_ENRL_ORIG_YR6_RT","FIRSTGEN_ENRL_ORIG_YR6_RT","NOT1STGEN_ENRL_ORIG_YR6_RT",
    "ENRL_ORIG_YR8_RT","LO_INC_ENRL_ORIG_YR8_RT","MD_INC_ENRL_ORIG_YR8_RT","HI_INC_ENRL_ORIG_YR8_RT","DEP_ENRL_ORIG_YR8_RT",
    "IND_ENRL_ORIG_YR8_RT","FEMALE_ENRL_ORIG_YR8_RT","MALE_ENRL_ORIG_YR8_RT","PELL_ENRL_ORIG_YR8_RT","NOPELL_ENRL_ORIG_YR8_RT",
    "LOAN_ENRL_ORIG_YR8_RT","NOLOAN_ENRL_ORIG_YR8_RT","FIRSTGEN_ENRL_ORIG_YR8_RT","NOT1STGEN_ENRL_ORIG_YR8_RT","C150_4_POOLED_SUPP",
    "PCTFLOAN","DEBT_MDN","GRAD_DEBT_MDN","WDRAW_DEBT_MDN","LO_INC_DEBT_MDN","MD_INC_DEBT_MDN","HI_INC_DEBT_MDN","DEP_DEBT_MDN",
    "IND_DEBT_MDN","PELL_DEBT_MDN","NOPELL_DEBT_MDN","CUML_DEBT_N","CUML_DEBT_P90","CUML_DEBT_P75","CUML_DEBT_P25",
    "CUML_DEBT_P10","loan_ever","DEBT_MDN_SUPP","GRAD_DEBT_MDN_SUPP","GRAD_DEBT_MDN10YR_SUPP","CDR2","CDR3","COMPL_RPY_1YR_RT",
    "NONCOM_RPY_1YR_RT","LO_INC_RPY_1YR_RT","MD_INC_RPY_1YR_RT","HI_INC_RPY_1YR_RT","DEP_RPY_1YR_RT","IND_RPY_1YR_RT",
    "PELL_RPY_1YR_RT","NOPELL_RPY_1YR_RT","FEMALE_RPY_1YR_RT","MALE_RPY_1YR_RT","FIRSTGEN_RPY_1YR_RT","NOTFIRSTGEN_RPY_1YR_RT",
    "RPY_3YR_RT","COMPL_RPY_3YR_RT","NONCOM_RPY_3YR_RT","LO_INC_RPY_3YR_RT","MD_INC_RPY_3YR_RT","HI_INC_RPY_3YR_RT","DEP_RPY_3YR_RT",
    "IND_RPY_3YR_RT","PELL_RPY_3YR_RT","NOPELL_RPY_3YR_RT","FEMALE_RPY_3YR_RT","MALE_RPY_3YR_RT","FIRSTGEN_RPY_3YR_RT",
    "NOTFIRSTGEN_RPY_3YR_RT","RPY_5YR_RT","COMPL_RPY_5YR_RT","NONCOM_RPY_5YR_RT","LO_INC_RPY_5YR_RT","MD_INC_RPY_5YR_RT",
    "HI_INC_RPY_5YR_RT","DEP_RPY_5YR_RT","IND_RPY_5YR_RT","PELL_RPY_5YR_RT","NOPELL_RPY_5YR_RT","FEMALE_RPY_5YR_RT",
    "MALE_RPY_5YR_RT","FIRSTGEN_RPY_5YR_RT","NOTFIRSTGEN_RPY_5YR_RT","RPY_7YR_RT","COMPL_RPY_7YR_RT","NONCOM_RPY_7YR_RT",
    "LO_INC_RPY_7YR_RT","MD_INC_RPY_7YR_RT","HI_INC_RPY_7YR_RT","DEP_RPY_7YR_RT","IND_RPY_7YR_RT","PELL_RPY_7YR_RT",
    "NOPELL_RPY_7YR_RT","FEMALE_RPY_7YR_RT","MALE_RPY_7YR_RT","FIRSTGEN_RPY_7YR_RT","NOTFIRSTGEN_RPY_7YR_RT","count_ed",
    "count_nwne_p6","count_wne_p6","mn_earn_wne_p6","md_earn_wne_p6","pct10_earn_wne_p6","pct25_earn_wne_p6",
    "pct75_earn_wne_p6","pct90_earn_wne_p6","sd_earn_wne_p6","gt_25k_p6","mn_earn_wne_inc1_p6","mn_earn_wne_inc2_p6",
    "mn_earn_wne_inc3_p6"
  )
# This dropped name set is part of the base query string...
fieldNames <- setdiff(fieldNames,c('CITY','LONGITUDE','LATITUDE','ZIP','region','LOCALE','CURROPER')) 
fieldNames <- grep('PCIP',fieldNames,value = TRUE,invert = TRUE) # drop the discipline fields; they're grabbed separately.

# Upper case variables are from 2013 and lower case variables are from the Treasury dataset of 2005, except 'region'.
# fromS10 <- grep('ENRL_ORIG',fieldNames)
fromS11 <- which(fieldNames %in% grep('^[A-Z]|region',fieldNames))
fromS05 <- setdiff(seq_along(fieldNames),fromS11)
dfNames <- ifelse(grepl('^[A-Z]|region',fieldNames),fieldNames,paste(fieldNames,"2005",sep="_")) # put a '_2005' suffix on the Treasury variables.

makeQuery <- function(year,fieldNames,dfNames) {
  paste("SELECT UNITID unitID,
        INSTNM College,
        CONTROL CollegeType,
        PREDDEG degree,
        CURROPER currop,
        DISTANCEONLY distance,
        RELAFFIL relaffil,
        CITY city,
        LONGITUDE lon,
        LATITUDE  lat,
        ZIP zip,
        st_fips state,
        region region,
        LOCALE locale,
        CCBASIC ccbasic,
        CCUGPROF ccugprof,
        Year Year,",
        paste(fieldNames,' ',dfNames,sep="",collapse = ","),
        "FROM Scorecard 
        WHERE Year=",year)
}

queryString2005 <- makeQuery(2005,fieldNames,dfNames)
queryString2013 <- makeQuery(2013,fieldNames,dfNames)


discNames <- gsub('^([A-Z][-a-z]+)[, and/]*([A-Z][a-z]+).*','\\1\\2',discgrp$LABEL)
discgrp %<>% tbl_df %>% mutate(discName = discNames)
queryStringDscplns <- makeQuery(2013,discgrp$VARIABLE.NAME,discNames)

# Retrieve the data for 2005 because that is the latest with Treasury data on student families.
student2005 <- dbGetQuery(db,queryString2005)
# Retrieve the data for 2013 because that's the latest data.
student2013 <- dbGetQuery(db,queryString2013)
disciplines2013 <- dbGetQuery(db,queryStringDscplns)

# Disconnect the database.
dbDisconnect(db)

# Join the years together.
student2013 <- student2013[!sapply(student2013,function(x) all(is.na(x)))] %>% tbl_df
student2005 <- student2005[!sapply(student2005,function(x) all(is.na(x)))] %>% tbl_df
# student <- student2013 %>% tbl_df %>% 
#   dplyr::select(-contains('_2005')) %>% 
#   inner_join(student2005 %>% dplyr::select(unitID,contains('_2005')), by = "unitID")
student <- student2013 %>% 
  inner_join(student2005 %>% dplyr::select(unitID,contains('_2005')), by = "unitID")
# Add the dominant disciplines
discplns <- disciplines2013 %>% tbl_df %>%
  filter(unitID %in% student$unitID) %>%
  mutate(irow = seq_len(nrow(.))) %>%
  gather(key=Discipline,value=Proportion,one_of(discNames)) %>%
  group_by(irow) %>%
  arrange(desc(Proportion)) %>%
  summarize(
    unitID   = first(unitID),
    College  = first(College),
    domDisc1 = first(Discipline),
    pctDisc1 = 100*first(Proportion),
    domDisc2 = nth(Discipline,n=2),
    pctDisc2 = 100*nth(Proportion,n=2)
  ) %>%
  mutate(
    isSpecialty = pctDisc1 > 70,
    isTheological = grepl('Theolo|Relig',domDisc1) | grepl('Theolo|Relig',domDisc2)
  )

# CALLED THE MAIN DATA TABLE `student` BECAUSE ORIGINALLY ONLY CONTAINED
# `student` CATEGORY OF DATA FIELDS. NOW THIS LEGACY NAME AWKWARDLY CONFUSES
# THIS DATA TABLE OF UNIVERSITIES/COLLEGES WITH THE OTHER DATA OBJECTS
# CONTAINING SPECIFIC STUDENT PROPERTIES -- SEE LATER...

# Now add the disciplines as grouped
student %<>% 
  inner_join(discplns %>% dplyr::select(-College),by='unitID')

usa.states.dc <- c("District of Columbia",state.name)
student0 <- student
```
***NOTE:*** At this point the foundational data table is `student0`, which is further
manipulated as `student`. This includes filtering out colleges we're not
interested in. See the printouts below to get a sense of what data were kept.
```{r}

# Filter down to just the colleges meeting the following criteria:
student  <- student0 %>% 
  filter(CollegeType != 'Private for-profit',
         currop      == 'Currently certified as operating',
         distance    == 'Not distance-education only',
         degree      == "Predominantly bachelor's-degree granting",
         as.character(region)      != 'U.S. Service Schools',
         !grepl('^Associate',ccbasic),
         state %in% usa.states.dc,
         !is.na(ccbasic),
         !is.na(UGDS),
         !is.na(pell_ever_2005),
         !isTheological,
         !is.na(mn_earn_wne_inc2_p6_2005))

# Turn character columns into factors.
student <- student %>% lapply(
  function(x) 
    if(is.character(x) | is.factor(x) | is.logical(x)) {
      if('PrivacySuppressed' %in% x) {
        tmp<-x;tmp[x=='PrivacySuppressed']<-NA;as.numeric(tmp)
      } else {
        factor(gsub("’","'",x))
      }
    } else as.numeric(x)) %>% 
  as.data.frame %>% tbl_df

# add a religious affiliation indicator variable:
student %<>% 
  mutate(relaffilFLAG=factor(ifelse(is.na(relaffil),
                                    FALSE,
                                    !grepl('^[Nn]ot ',as.character(relaffil)))))


a<- student %>% dplyr::select(grep('^(SAT|ACT)',names(.))) %>%  apply(1,function(x) !all(is.na(x)))
student %<>% mutate(stdzdtest=a)

show(dim(student))
#show(summary(student))

```
Drop all the Penn St. and U. of Pitt. schools except the main campus because
same 2005 Treasury data entered in all rows for each, resp.
```{r}
stdttmp <- student %>% filter((College %in% c('Pennsylvania State University-Main Campus',
                                              'University of Pittsburgh-Pittsburgh Campus')) |
                                !(College %in% grep('Pennsylvania State University|University of Pittsburgh',
                                                    College,value=TRUE)))
show(setdiff(student$College,stdttmp$College))
show(grep('Pennsylvania State University|University of Pittsburgh',stdttmp$College,value=TRUE))
student <- stdttmp
rm(stdttmp)
student <- student[!sapply(student,function(x) length(which(is.na(x)))>1000)]
# Drop schools without earnings data:
student %>% filter(is.na(pct10_earn_wne_p6_2005) | pct10_earn_wne_p6_2005==0) %>% 
  dplyr::select(unitID,College,UGDS,matches('pct.+earn')) %>% 
  print(n=60)
student %<>% 
  filter(!is.na(pct10_earn_wne_p6_2005) & pct10_earn_wne_p6_2005>0) 

#show(summary(student))
```
Most selective schools according to admissions rate:
```{r}
# Most selective schools according to admissions rate:
# student %>% 
#   filter(!is.na(ADM_RATE) & ADM_RATE<0.2 & UGDS>600) %>% 
#   dplyr::select(unitID,College,UGDS,matches('SAT|ACT'),ADM_RATE) %>% 
#   arrange(ADM_RATE) %>% 
#   print(n=40)
student %>% 
  filter(!is.na(ADM_RATE) & ADM_RATE<0.2 & UGDS>600) %>% 
  dplyr::select(unitID,College,UGDS,contains('Disc'),ADM_RATE,pell_ever_2005) %>% 
  arrange(ADM_RATE) %>% 
  print(n=40)
#===================================================
```

## Building Conditional Probabilities & Bayes Factors
What follows are code chunks to generate specific contextual data tables
of conditional probabilities and Bayes factors for different aspects of the
college experience for a student at each college, as was described above. 
Note that I fudged the Bayes factors -- even after defining the functions to
compute them properly. Also, I did a bit of exploration with poLCA, but
commented it out. (Using ggplot2, it's easy to visually explore clustering
solutions using 'small multiples' (i.e., facet_wrap). But not shown here....)

### Ethnicity:
```{r}
ethnicity <- student %>% 
  dplyr::select(unitID,College,contains("UGDS")) %>%
  mutate(probSchool = UGDS/sum(UGDS),
         totprob    = rowSums(as.matrix(.[4:ncol(.)])))
ethnicity %>% print(n=20)
```
This computes P(E | not(H)) the conditional probability of evidence given the
hypothesis is FALSE using inputs of cpH = P(E | H) probability of evidence
given the hypothesis is TRUE and pH = P(H) the marginal (prior) probability
of hypothesis being TRUE. But, in this case of almost 1500 schools, where "H"
is school si, P(E|not(H)) is approximately P(E) = sum(P(E|H)*P(H);over H). 
So, could skip this function and just use P(E).

***IMPORTANT:*** I use log10 of the Bayes factors, which is a bit unusual,
but I know powers of 10 better than powers of 2 or *e*.
```{r}
cpNotH <- function(cpH,pH){
  pEH <- cpH * pH
  pE  <- sum(pEH)
  return((pE - pEH)/(1 - pH))
}
BFactor <- function(cpH,cpNotH,logFunc=log10){
  x <- cpH / cpNotH
  return(if(is.function(logFunc)) logFunc(x) else x)
}
makeCpNotH <- function(x) mapply(cpNotH,cpH=x,pH=ethnicity$probSchool)
makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*ethnicity$probSchool,na.rm=TRUE))} # Approximate

ethnicBF <- ethnicity %>% 
  dplyr::select(contains("UGDS_")) %>% 
  mutate_each(funs(makeBF)) %>% 
  setNames(gsub("UGDS","BF",names(.))) %$%
  bind_cols(ethnicity[1:2],.) %>%
  mutate(BF_female_2005 = makeBF(student$female_2005))
#=====================================================
```
### Income:
```{r}
income <- student %>% 
  dplyr::select(unitID,College,UGDS,DEP_STAT_PCT_IND,contains("INC_PCT")) %>%
  mutate(probSchool = UGDS/sum(UGDS),
         totprob      = rowSums(as.matrix(.[c("INC_PCT_LO","INC_PCT_M1","INC_PCT_M2","INC_PCT_H1",
                                              "INC_PCT_H2")])),
         totprobdep   = rowSums(as.matrix(.[c("DEP_INC_PCT_LO","DEP_INC_PCT_M1","DEP_INC_PCT_M2",
                                              "DEP_INC_PCT_H1","DEP_INC_PCT_H2")])),
         totprobind   = rowSums(as.matrix(.[c("IND_INC_PCT_LO","IND_INC_PCT_M1","IND_INC_PCT_M2",
                                              "IND_INC_PCT_H1","IND_INC_PCT_H2")])),
         totprob2dep  = rowSums(as.matrix(.[c("DEP_INC_PCT_LO","DEP_INC_PCT_H2")])),
         totprob2ind  = rowSums(as.matrix(.[c("IND_INC_PCT_LO","IND_INC_PCT_H2")])))
income %>% print(n=20)

# To avoid dropping many schools due to missing values will hold 7 variables
# for income: DEP_STAT_PCT_IND, DEP_INC_PCT_LO, DEP_INC_PCT_H2 &
# DEP_INC_GTLO_LTH2, IND_INC_PCT_LO, IND_INC_PCT_H2 & IND_INC_GTLO_LTH2; where
# the x_INC_GTLO_LTH2 variables are 1 - x_INC_PCT_H2 - x_INC_PCT_LO; and the
# sum of the latter 6 columns = 1. So must mutate the current IND_... columns
# by multiplying by DEP_STAT_PCT_IND. An "x" is appended to the end of the
# newly computed variable names.

income %<>% mutate(probSchool = UGDS/sum(UGDS),
                   DEP_STAT_PCT_INDx  = DEP_STAT_PCT_IND,
                   DEP_STAT_PCT_DEPx  = 1 - DEP_STAT_PCT_INDx,
                   DEP_INC_PCT_LOx    = DEP_INC_PCT_LO*DEP_STAT_PCT_DEPx,
                   DEP_INC_PCT_HI2x   = DEP_INC_PCT_H2*DEP_STAT_PCT_DEPx,
                   DEP_INC_GTLO_LTH2x = DEP_STAT_PCT_DEPx - DEP_INC_PCT_LOx - DEP_INC_PCT_HI2x,
                   IND_INC_PCT_LOx    = IND_INC_PCT_LO*DEP_STAT_PCT_INDx,
                   IND_INC_PCT_HI2x   = IND_INC_PCT_H2*DEP_STAT_PCT_INDx,
                   IND_INC_GTLO_LTH2x = DEP_STAT_PCT_IND - IND_INC_PCT_LOx - IND_INC_PCT_HI2x,
                   totprobx = DEP_INC_PCT_LOx + DEP_INC_PCT_HI2x + DEP_INC_GTLO_LTH2x + IND_INC_PCT_LOx + IND_INC_PCT_HI2x + IND_INC_GTLO_LTH2x) %>%
  dplyr::select(unitID,College,UGDS,probSchool,matches("x$"))
income %>% print(n=20)
#income[income$College=='California Institute of Technology',
#        c('IND_INC_PCT_LOx','IND_INC_PCT_HI2x','IND_INC_GTLO_LTH2x','totprobx')] <- c(0,0,0,1)

makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*income$probSchool,na.rm=TRUE))} # Approximate
incomeBF <- income %>% 
  dplyr::select(matches("x$"),-totprobx) %>% 
  mutate_each(funs(makeBF)) %>% 
  setNames(gsub("(.+)x$","BF_\\1x",names(.))) %$%
  bind_cols(income[1:4],.)

incomeBFdiscrete <- incomeBF %>% dplyr::select(5:ncol(incomeBF)) %>% filter(complete.cases(.)) %>%
  lapply(function(x) discretize(data.frame(x[ifelse(is.na(x),FALSE,x>-5)]),method='quantile',breaks=5))
incomeBFlowestState <- incomeBF %>% dplyr::select(5:ncol(incomeBF)) %>% filter(complete.cases(.)) %>%
  lapply(function(x) ifelse(x<= -5,"(-Inf,-5]",'FALSE'))
incomeBFd <- mapply(function(x,y) {
  tmp <- x
  tmp[tmp=='FALSE'] <- as.character(y[[1]])
  return(factor(tmp,levels=c("(-Inf,-5]",levels(y[[1]]))))},
  incomeBFlowestState,incomeBFdiscrete,SIMPLIFY = FALSE)
names(incomeBFd) <- names(incomeBFdiscrete)
incomeBFd %<>% data.frame %>% tbl_df
rm(list=c('incomeBFdiscrete','incomeBFlowestState'))

incomeBFd <-  incomeBF %>% 
  filter(complete.cases(incomeBF %>% dplyr::select(5:ncol(incomeBF)))) %>% 
  dplyr::select(1:4) %>% 
  bind_cols(incomeBFd)
```

Here, I did some cluster analysis using package `poLCA`  but ultimately set
it aside to be dug into later...
```{r}
# #-------------------------------------
# bmod <- poLCA(as.formula(paste0('cbind(',paste(names(incomeBFd)[-(1:4)],collapse=','),') ~ 1')),
#               data=incomeBFd,nclass = 9,nrep=13,na.rm=FALSE)
# new.probs.start <- poLCA.reorder(bmod$probs.start,order(bmod$P,decreasing=TRUE))
# bmod <- poLCA(as.formula(paste0('cbind(',paste(names(incomeBFd)[-(1:4)],collapse=','),') ~ 1')),
#               data=incomeBFd,nclass = 9,probs.start=new.probs.start,na.rm=FALSE)
# 
# # Compute the mutual information shared by each variable with the class labels:
# miBFclass <- incomeBFd %>% 
#   mutate(class=bmod$predclass) %$% 
#   sapply(grep('^BF',names(.),value=TRUE),mInfo,ynm='class',data=.) %>% 
#   sort(decreasing=TRUE) %T>%
#   print
# incomeBFd %>% 
#   mutate(class=bmod$predclass) %>% 
#   inner_join(incomeBF,by='unitID') %>% 
#   group_by(class) %>% 
#   summarize_each(funs(median),matches('x.y$')) %>%
#   dplyr::select(one_of(names(miBFclass))) %>% 
#   print(width=Inf)
# 
# incomeBFd %>% 
#   dplyr::select(-c(1,3:4)) %>% 
#   mutate(class=bmod$predclass) %>% 
#   filter(class==9) %>% 
#   print(n=20,width=Inf)
# 
# incomeBFd %>% 
#   mutate(class=bmod$predclass) %>% 
#   filter(grepl("Cali.+Insti.+Tech|Harvard|Mass.+Inst.+Tech|Northwestern Univ|Princeton|Cornell U",
#                College)) %>%
#   dplyr::select(one_of(setdiff(names(.),names(miBFclass))),one_of(names(miBFclass))) %>% 
#   arrange(class) %>%
#   print(width=Inf)
# #---------------------------------------
#===================================================
```
### Student Aid:
```{r}
aid <- student %>% 
  dplyr::select(unitID,College,UGDS,matches("pell_ever|fsend")) %>%
  mutate(probSchool = UGDS/sum(UGDS),
         totprob    = rowSums(as.matrix(.[grepl('fsend',names(.))])))
aid %>% print(n=20)
makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*aid$probSchool,na.rm=TRUE))} # Approximate
aidBF <- aid %>% 
  dplyr::select(matches("pell_ever|fsend")) %>% 
  mutate_each(funs(makeBF)) %>% 
  setNames(paste0("BF_",names(.))) %$%
  bind_cols(aid[1:3],.)

g <- aidBF %>% gather(key=Variable,value=BF,-(1:3)) %>%
  ggplot(aes(x=BF,fill=Variable))
g + geom_density(alpha=0.3) + 
  facet_wrap(~Variable) + 
  scale_x_continuous(lim=c(-1,1)) + 
  theme(text=element_text(size=14,face = 'bold'))
#===================================================
```
### Earnings after graduation (6 years on)
```{r}
earnRepay <- student %>% 
  dplyr::select(unitID,College,UGDS,matches("^(RPY|CDR|.+earn_wne_p6)")) %T>% 
  print(n=20) %>%
  mutate(probSchool = UGDS/sum(UGDS),
         sdlog         = sqrt(log((sd_earn_wne_p6_2005/mn_earn_wne_p6_2005)^2+1)),
         meanlog       = log(mn_earn_wne_p6_2005^2/sqrt(mn_earn_wne_p6_2005^2+sd_earn_wne_p6_2005^2)),
         p_le30K       = plnorm(30.0E3,meanlog,sdlog),
         p_gt30Kle48K  = plnorm(48.0E3,meanlog,sdlog)  - p_le30K,
         p_gt48Kle75K  = plnorm(75.0E3,meanlog,sdlog)  - plnorm(48.0E3,meanlog,sdlog),
         p_gt75Kle110K = plnorm(110.0E3,meanlog,sdlog) - plnorm(75.0E3,meanlog,sdlog),
         p_gt110K      = plnorm(110.0E3,meanlog,sdlog,lower.tail=FALSE),
         totprob       = p_le30K + p_gt30Kle48K + p_gt48Kle75K + p_gt75Kle110K + p_gt110K) %>%
  dplyr::select(1:7,probSchool,starts_with('p_')) %T>%
  print(n=20)

makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*earnRepay$probSchool,na.rm=TRUE))} # Approximate
earnRepayBF <- earnRepay %>% 
  dplyr::select(-c(1:3,5:8)) %>% 
  mutate_each(funs(makeBF)) %>% 
  setNames(paste0("BF_",names(.))) %$%
  bind_cols(earnRepay[c(1:3,8)],.) %>%
  filter(complete.cases(.))

earnRepayBF %>% 
  dplyr::select(-3) %>% 
  inner_join(student %>% dplyr::select(unitID,ADM_RATE),by='unitID') %>% 
  filter(!is.na(ADM_RATE)) %>% 
  mutate(ADM_RATEd=discretize(.['ADM_RATE'],method='interval',breaks=6)[['ADM_RATE']]) %>% 
  arrange(desc(BF_p_gt110K)) %>% 
  filter(BF_p_le30K<0.5 & BF_p_gt110K>0.5) %$% 
  qplot(x=BF_p_le30K,y=BF_p_gt110K,color=ADM_RATEd) + geom_text(aes(label=College))
#===================================================
```
### College Setting: locale & region.
```{r}
student %$% { locale %>% summary %>% print }
student %$% { region %>% summary %>% print }

localeNames <- sort(unique(as.character(student$locale)))
localeAggregates <- setNames(c(gsub('([^ ]+) ([^ ]+).+','\\1\\2',localeNames[1:3]),
                               rep('Rural',3),
                               gsub('([^ ]+) ([^ ]+).+','\\1\\2',localeNames[7]),
                               rep('Suburb:Small/Midsize & Town:Fringe',2),
                               gsub('([^ ]+) ([^ ]+).+','\\1\\2',localeNames[10]),
                               'Suburb:Small/Midsize & Town:Fringe',
                               gsub('([^ ]+) ([^ ]+).+','\\1\\2',localeNames[12])),localeNames)
student %<>% mutate(localeAgg = factor(localeAggregates[as.character(locale)],
                                       levels=c('Rural','Town:Remote','Town:Distant','Suburb:Small/Midsize & Town:Fringe',
                                                'Suburb:Large','City:Small','City:Midsize','City:Large')))
makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*settingBF$probSchool,na.rm=TRUE))} # Approximate
settingBF <- student %>% 
  dplyr::select(unitID,College,UGDS,localeAgg,region) %>%
  mutate(probSchool = UGDS/sum(UGDS))
a <- settingBF %$% model.matrix(~localeAgg - 1,data=.)
colnames(a) <- gsub('^([^:]+):*(.+)$','BF_\\1\\2',colnames(a))
a %<>% as.data.frame %>% tbl_df %>% mutate_each(funs(makeBF))
settingBF %<>% dplyr::select(-localeAgg) %>% bind_cols(a)

a <- settingBF %>%
  mutate(regionName = factor(gsub(' ','',as.character(student$region)))) %$% 
  model.matrix(~regionName - 1,data=.)
colnames(a) <- gsub('^regionName','BF_',colnames(a))
a %<>% as.data.frame %>% tbl_df %>% mutate_each(funs(makeBF))
settingBF %<>% dplyr::select(-region) %>% bind_cols(a)
```
Didn't get around to adding in costs...
```{r}
#===================================================
#costBF
#===================================================
```
### Academics: Completion rates, admissions rates and SAT scores.
Took a look at scaled test scores, but ultimately settled upon the log-normal
fits to SATs below.
```{r}
# scaleSAT <- function(Score) (Score - 280)/(800 - 280)
# scaleACT <- function(Score) (Score -   9)/( 36 -   9)
# Impute missing values with the means
#lmtest <- lm(ACTCMMID ~ SAT_AVG,data=student)

# academicsBF <- student %>% 
#   filter(!(isSpecialty=='TRUE' & is.na(ADM_RATE))) %>%
#   dplyr::select(unitID,College,ADM_RATE,matches('(^C150)|(SAT|ACT)')) %>%
#   filter(!is.na(C150_4_POOLED_SUPP)) %>%
#   mutate(ADM_RATE = ifelse(is.na(ADM_RATE),mean(ADM_RATE,na.rm=TRUE),ADM_RATE)) %>%
#   mutate(SAT_AVG_scaled  = scaleSAT(ifelse(is.na(SAT_AVG) , mean(SAT_AVG,na.rm=TRUE), SAT_AVG))) %>%
#   mutate(ACTCMMID_scaled = scaleACT(ifelse(is.na(ACTCMMID),mean(ACTCMMID,na.rm=TRUE),ACTCMMID))) %>%
#   dplyr::select(1:3,ncol(.) - (2:0))

# a <- academicsBF %>% 
#   mutate_each(funs(scaleSAT),starts_with('SAT'),-ends_with('scaled')) %>% 
#   mutate_each(funs(scaleACT),starts_with('ACT'),-ends_with('scaled')) %>%
#   gather(key=Test,value=scaledScore,-unitID,-College,-ADM_RATE)
```
***NOTE:*** This is a bit of a kludge, but it results in a distribution for SAT
Verbal+Math scores for each school, from which Bayes factors for 200-pt
intervals are computed. (Kludginess: Adding the quartiles of Verbal & Math
DOES NOT equal the quartiles of Verbal+Math.)
```{r}
student %>% 
  mutate(SAT_25=SATVR25+SATMT25,SAT_75=SATVR75+SATMT75) %>% 
  filter(!is.na(SAT_AVG) & !is.na(SAT_75)) %>% 
  ggplot(aes(x=SAT_AVG)) + geom_density(fill='orange',alpha=0.3) + 
  geom_density(aes(x=SAT_25),fill='red',alpha=0.1) + 
  geom_density(aes(x=SAT_75),fill='green',alpha=0.1)

academics <- student %>% 
  mutate(SAT_25=SATVR25+SATMT25,SAT_75=SATVR75+SATMT75) %>% 
  filter(!is.na(SAT_AVG) & !is.na(SAT_75)) %>% 
  dplyr::select(unitID,College,UGDS,ADM_RATE,C150_4_POOLED_SUPP,SAT_25,SAT_AVG,SAT_75)


# Clumsily fit log-normal distributions to the SAT score quartiles for each school.
fr <- function(x,SAT_AVG,SAT_25,SAT_75,probs=c(0.25,0.75)) {
  sdl <- x[1]
  q <- qlnorm(p=probs,meanlog=log(SAT_AVG) - sdl^2/2,sdlog=sdl)
  log(SAT_25/q[1])^2 + log(SAT_75/q[2])^2
}
getMuSd <- function(SAT_AVG,SAT_25,SAT_75) {
  probs  <- c(0.25,0.75)
  SAT_AVG0 <- SAT_AVG
  if(SAT_75==1600) {
    probs   <- probs/(0.75+1.0E-5)
    SAT_AVG <- (SAT_AVG0 - 1600*(1-(0.75+1.0E-5)))/(0.75+1.0E-5)
  }
  soln <- optim(c(0.05), fr,lower=1.0E-3,upper=0.2,method='L-BFGS-B',
                SAT_AVG=SAT_AVG,SAT_25=SAT_25,SAT_75=SAT_75,probs=probs)
  sdl  <- soln$par
  meanlog  <- log(SAT_AVG) - sdl^2/2
  pSAT_AVG <- exp(meanlog+0.5*sdl^2)
  q   <- qlnorm(p=probs,meanlog=meanlog,sdlog=sdl)
  if(SAT_75==1600){
    pSAT_AVG <- pSAT_AVG*(0.75+1.0E-5) + 1600*(1-(0.75+1.0E-5))
  }
  return(c(meanlog=meanlog,sdlog=sdl,pSAT_25=q[1],pSAT_AVG=pSAT_AVG,pSAT_75=q[2]))
}
```
Get the log-normal parameters of the SAT distribution for each school
```{r}
academics <- academics %$% {bind_cols(.,as.data.frame(t(mapply(getMuSd,SAT_AVG,SAT_25,SAT_75))))}
academics %>% 
  filter(grepl('Harvard|Cal.+Inst.+Tech|Mass.+Inst.+Tech|Princeton U|Northwestern U|Cornell U',College)) %>% 
  arrange(desc(SAT_AVG))
```
Discretize the SAT distributions
```{r}
academics %<>%
  filter(!is.na(C150_4_POOLED_SUPP)) %>%
  mutate(probSchool = UGDS/sum(UGDS),
         p_le800        = plnorm( 800,meanlog,sdlog),
         p_gt800le1000  = plnorm(1000,meanlog,sdlog) - p_le800,
         p_gt1000le1200 = plnorm(1200,meanlog,sdlog) - plnorm(1000,meanlog,sdlog),
         p_gt1200le1400 = plnorm(1400,meanlog,sdlog) - plnorm(1200,meanlog,sdlog),
         p_gt1400       = plnorm(1400,meanlog,sdlog,lower.tail=FALSE),
         totprob        = p_le800 + p_gt800le1000 + p_gt1000le1200 + p_gt1200le1400 + p_gt1400) %>%
  print
```
Compute Bayes Factors for the discretized SAT score distributions of each school
```{r}
makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*academics$probSchool,na.rm=TRUE))} # Approximate
academicsBF <- academics %>% 
  dplyr::select(-c(1:3,matches('SAT|log|prob'))) %>% 
  mutate_each(funs(makeBF)) %>% 
  setNames(gsub("^p","BF_SAT",names(.))) %>%
  setNames(gsub("^([^B])","BF_\\1",names(.))) %$%
  bind_cols(academics[1:3],.)

academicsBF %>% 
  filter(grepl('Harvard|Cal.+Inst.+Tech|Mass.+Inst.+Tech|Princeton U|Northwestern U|Cornell U',College)) %>% 
  arrange(desc(BF_SAT_gt1400)) %>% 
  print(width=Inf)
```
Now add in the disciplines:
```{r}
makeBF <- function(x) {x <- 1.0E-9 + x;log10(x/sum(x*academics$probSchool,na.rm=TRUE))} # Approximate
discplnsBF <-  academics %>% dplyr::select(unitID,probSchool) %>% inner_join(disciplines2013,by='unitID') %>%
  mutate_each(funs(makeBF),-seq_len(which(names(.) == discNames[1])-1)) %>% 
  setNames(c(setdiff(names(.),discNames),paste0('BF_',discNames)))

academicsBF %<>% inner_join(discplnsBF %>% dplyr::select(unitID,starts_with('BF_')),by='unitID')
```
## Bringing the Data Together
At this point, we've got enough pieces of the puzzle to construct a working
model that demonstrates the concept of a decision analysis tool that takes as
input a student's demographics and behavioral traits and the dataset of
college Bayes factors and outputs a list of the top-N colleges maximizing the
utility for that specific student.

First join the separate data tables of Bayes factors into a single table.
```{r}

# NOW: Build the model covariate data frame with all possible covariate
# candidates. Each student's model only uses a subset of these columns.  See
# function 'utility'....
studentBF <- student     %>% dplyr::select(unitID,College,UGDS,ADM_RATE) %>%
  inner_join(incomeBF    %>% dplyr::select(-College,-UGDS), by='unitID') %>%
  inner_join(ethnicBF    %>% dplyr::select(-College),       by='unitID') %>%
  inner_join(aidBF       %>% dplyr::select(-College,-UGDS), by='unitID') %>%
  inner_join(settingBF   %>% dplyr::select(-College,-UGDS,-probSchool), by='unitID') %>%
  inner_join(earnRepayBF %>% dplyr::select(-College,-UGDS,-probSchool), by='unitID') %>%
  #  inner_join(costBF      %>% dplyr::select(-College,-UGDS), by='unitID') %>%
  inner_join(academicsBF %>% dplyr::select(-College,-UGDS), by='unitID') %>%
  mutate(BF_prior = log10(nrow(.)*probSchool)) %>%
  print(n=20)
```
## Mapping Student Traits to College Properties
Get the coefficients for the student profile:
Two levels of coefficients: 1.Property Category level, 2. Subcategory level

Trait level has 4 types with impact on properties as follows when each trait increases from low to medium to high: 

1. Risk      -- willingness to surround self with people & settings very different from self & origins;
    + Weights homogeneity & sameness as self lower than other traits.
2. Vision    -- willingness to look past short-term suitability in pursuit of future;
    + Weights completion rates, future earnings, debt & debt repayment rates higher than current campus setting compatability.
3. Breadth   -- willingness to entertain a variety of academic disciplines;
    + Weights predominance of own major discipline lower than otherwise. 
4. Challenge -- willingness to embrace highest academic rigor
    + Weights admissions selectiveness, high SAT & higher than otherwise.

                
Could also add in more or less uncertainty to beta's according to traits as well...

Traits Have Impact On Four Property Categories Listed in Rows:

|                              |      Risk       |      Vision       |      Breadth       |      Challenge       |      Properties
|  ----------------------------|-----------------|-------------------|--------------------|----------------------|---------------------------
|   1. Self & Origins          | same to differ  |      ------       |      -------       |        ----          | ethn,inc,1stgen,usbrn,pov
|   2. Campus Setting & Folks  |homog to diverse |      ------       |      -------       |        ----          | reg,loc,siz
|   3. Academics               |      ----       |      ------       |  narrow to broad   | low to high sat/adm  | sat,admrt,disc
|   4. After-College Prospects |      ----       |low to high dbt/ern|      -------       |        ----          | comprt,earn,rpy,dfltrt

```{r}
# This maps the student profile property names to the college Bayes factor data table ('studentBF') column names.

propertyMap <- list(
  ethnicity  = c(white='BF_WHITE',black='BF_BLACK',hispanic='BF_HISP',asian='BF_ASIAN'),
  income     = list(dependent   = c(lt30K="BF_DEP_INC_PCT_LOx",gt30Kle110K="BF_DEP_INC_GTLO_LTH2x",gt110K="BF_DEP_INC_PCT_HI2x"),
                    independent = c(lt30K="BF_IND_INC_PCT_LOx",gt30Kle110K="BF_IND_INC_GTLO_LTH2x",gt110K="BF_IND_INC_PCT_HI2x")),
  sat        = c(le800='BF_SAT_le800',  gt800le1000='BF_SAT_gt800le1000', gt1000le1200='BF_SAT_gt1000le1200', 
                 gt1200le1400='BF_SAT_gt1200le1400', gt1400='BF_SAT_gt1400'),
  discipline = setNames(paste0('BF_',discNames),discNames),
  fasfa      = c(fsend_1='BF_fsend_1_2005',fsend_2='BF_fsend_2_2005',fsend_3='BF_fsend_3_2005',fsend_4='BF_fsend_4_2005',fsend_5='BF_fsend_5_2005'),
  region     = c(FarWest='BF_FarWest(AK,CA,HI,NV,OR,WA)',GreatLakes='BF_GreatLakes(IL,IN,MI,OH,WI)',MidEast='BF_MidEast(DE,DC,MD,NJ,NY,PA)',
                 NewEngland='BF_NewEngland(CT,ME,MA,NH,RI,VT)',Plains='BF_Plains(IA,KS,MN,MO,NE,ND,SD)',RockyMountains='BF_RockyMountains(CO,ID,MT,UT,WY)',
                 Southeast='BF_Southeast(AL,AR,FL,GA,KY,LA,MS,NC,SC,TN,VA,WV)',Southwest='BF_Southwest(AZ,NM,OK,TX)'),
  locale     = c(Rural='BF_localeAggRural',TownRemote='BF_localeAggTownRemote',TownDistant='BF_localeAggTownDistant',
                 SuburbSmallMid='BF_localeAggSuburbSmall/Midsize & Town:Fringe',SuburbLarge='BF_localeAggSuburbLarge',
                 CitySmall='BF_localeAggCitySmall',CityMidsize='BF_localeAggCityMidsize',CityLarge='BF_localeAggCityLarge'),
  earnings   = c(le30K='BF_p_le30K',  gt30Kle48K='BF_p_gt30Kle48K', gt48Kle75K='BF_p_gt48Kle75K', gt75Kle110K='BF_p_gt75Kle110K', gt110K='BF_p_gt110K' ),
  prior      = 'BF_prior',
  completion = 'BF_C150_4_POOLED_SUPP',
  admission  = 'BF_ADM_RATE',
  gender     = c(female='BF_female_2005',male='BF_female_2005')
)
```
## Model Parameters -- The Partworths of the Utility Function
A critical determinant of how well the tool works is the capturing of
specific student value assessments of college properties. This is done using
parameters beta in the utility function.  The function below does this.

This is just to show one of the many ways in which this can be done.
```{r}
getParameters <- function(studentProfile,propertyMap){
  # Get the coefficients for the student profile:
  propCategory <- c('SelfOrigin','Setting','Academics','Prospects')
  traitLvls    <- setNames(as.list(studentProfile$traits),c('Risk','Vision','Breadth','Challenge'))
  propsByTrait <- list(Risk      = list(SelfOrigin = c('ethnicity','income'),
                                        Setting    = c('region','locale')), 
                       Vision    = list(Prospects  = c('completion','earnings'),
                                        Setting    = c('region','locale')), 
                       Breadth   = list(Academics  = c('discipline')), 
                       Challenge = list(Academics  = c('sat','admission','discipline')))
  
  # beta <- matrix(rep(1,length(xBF)),ncol=1,dimnames=list(c('ethnicity','income','sat','fasfa'),'Utility'))
  beta <- matrix(rep(1,length(propertyMap)),ncol=1)
  rownames(beta) <- names(propertyMap)
  beta['admission',1] <- -1.0  # prefer schools that have lower admissions rates
  beta['locale'   ,1] <-  0.5  # downweight locale
  beta['fasfa'    ,1] <-  0.3  # downweight fasfa
  # assumes stronger influence of gender for women than men...
  beta['gender'   ,1] <- ifelse(studentProfile$gender == 'female',0.5,-0.1) 
  # (^_^) ...there's gotta be a better way of doing this...
  for(pCat in propCategory){
    prpnm <- propsByTrait$Risk[[pCat]]
    if (!is.null(prpnm)){
      # must have vectors for each trait level with as many elements as are in the vectors of the propsByTrait list
      beta[prpnm,1] <- beta[prpnm,1] * switch(traitLvls$Risk,
                                              L=list(SelfOrigin=c(ethnicity=3.0,income=2.0),
                                                     Setting=c(region=3.0,locale=3.0))[[pCat]],
                                              M=list(SelfOrigin=c(ethnicity=0.3,income=0.3),
                                                     Setting=c(region=1.0,locale=1.0))[[pCat]],
                                              H=list(SelfOrigin=c(ethnicity=0.1,income=0.1),
                                                     Setting=c(region=0.1,locale=0.1))[[pCat]])
    }
    prpnm <- propsByTrait$Vision[[pCat]]
    if (!is.null(prpnm)){
      beta[prpnm,1] <- beta[prpnm,1] * switch(traitLvls$Vision,
                                              L=list(Prospects=c(completion=0.1,earnings=0.1),
                                                     Setting  = c(region=1.5,locale=1.5))[[pCat]],
                                              M=list(Prospects=c(completion=1.0,earnings=0.5),
                                                     Setting  = c(region=1.0,locale=1.0))[[pCat]],
                                              H=list(Prospects=c(completion=3.0,earnings=3.0),
                                                     Setting  = c(region=0.5,locale=0.5))[[pCat]])
    }
    prpnm <- propsByTrait$Breadth[[pCat]]
    if (!is.null(prpnm)){
      beta[prpnm,1] <- beta[prpnm,1] * switch(traitLvls$Breadth,
                                              L=list(Academics=c(discipline=3.0))[[pCat]],
                                              M=list(Academics=c(discipline=1.0))[[pCat]],
                                              H=list(Academics=c(discipline=0.3))[[pCat]])
    }
    prpnm <- propsByTrait$Challenge[[pCat]]
    if (!is.null(prpnm)){
      beta[prpnm,1] <- beta[prpnm,1] * switch(traitLvls$Challenge,
                                              L=list(Academics=c(sat=0.1,admission=0.1,discipline=0.3))[[pCat]],
                                              M=list(Academics=c(sat=1.0,admission=1.0,discipline=1.0))[[pCat]],
                                              H=list(Academics=c(sat=3.0,admission=3.0,discipline=3.0))[[pCat]])
    }
  }
  
  # Also, make the beta's of traits sum to 1 at each level so that final utility
  # is roughly same magnitude as a typical Bayes factor for school properties.
  # This facilitates interpretation of utility in similar fashion as a raw Bayes factor.
  beta[,1] <- beta[,1]/sum(beta[,1])
  
  return(beta)
}


```
## Utility Function to be Maximized for the Student
Now the utility function of the discrete choice model:
```{r}
#===================================================================
# Here's the Decision Analysis Model: Probabilistic, Utility-Based Discrete Choice Model
utility <- function(studentBF,studentProfile,propertyMap,dump.covariates=FALSE){
  # studentBF must come in as a 1-row matrix!!!
  
  covarnames <- sapply(names(propertyMap),function(propnm){
    if(propnm == 'income'){
      return(propertyMap$income[[ifelse(studentProfile$dependent,
                                        'dependent',
                                        'independent')]][studentProfile[[propnm]]])
    }
    levelName <- studentProfile[[propnm]]
    if(is.null(levelName)) return(propertyMap[[propnm]]) 
    return(propertyMap[[propnm]][levelName])
  })
  if(is.null(dim(studentBF))) {
    xBF <- matrix(studentBF[covarnames],nrow=1,dimnames=list(NULL,covarnames))
  } else {
    xBF <- matrix(studentBF[,covarnames],nrow=1,dimnames=list(NULL,covarnames))
  }
  
  # Assign useful column names...
  colnames(xBF) <- paste(colnames(xBF),names(covarnames),sep=':')
  
  #show(xBF)
  
  # Compute the utility the student derives from the school.
  beta <- setNames(studentProfile$beta,colnames(xBF))
  u    <- xBF %*% beta
  
  if (dump.covariates){
    wxBF <- xBF * matrix(beta,nrow=nrow(xBF),ncol=ncol(xBF),byrow = TRUE)
    colnames(wxBF) <- colnames(xBF)
    return(wxBF)
  } else {
    return(u[1])
  }
}
```
## Archetypical Student Profiles
We now specify student profiles and test out the model to judge how
reasonable are its rankings of colleges per specific student criteria.
```{r}

#=====================================================
# Define (3 levels)^(4 traits) = 3^4 = 81 student trait archetypes:
# Trait level has 4 categories with impact on subtraits as follows when each
# trait increases from low to medium to high:
# 1. Risk      -- willingness to surround self with people & settings very different from self & origins;
#                 i. Weights homogeneity & sameness as self lower than other traits.
# 2. Vision    -- willingness to look past short-term suitability in pursuit of future;
#                 i. Weights completion rates, future earnings, debt & debt repayment rates higher 
#                    than current campus setting compatability.
# 3. Breadth   -- willingness to entertain a variety of academic disciplines;
#                 i. Weights predominance of own major discipline lower than otherwise. 
# 4. Challenge -- willingness to embrace highest academic rigor
#                 i. Weights admissions selectiveness, high SAT & higher than otherwise.
makeTraitLabels <- function(trait,ntrait,levels){
  # Note that this must be called with positive integers for 'trait' & 'ntrait' and ntrait <= length(levels) 
  # (with intention to use ntrait==length(levels)) where 'levels' is a list of character vectors.
  # Will return a character vector with length = prod(sapply(levels,length)).
  if(trait>=ntrait) return(levels[[trait]])
  return(c(sapply(levels[[trait]],paste0,makeTraitLabels(trait+1,ntrait,levels))))
}
traitLevels  <- list(Risk=c('Lr','Mr','Hr'),Vision=c('Lv','Mv','Hv'),Breadth=c('Lb','Mb','Hb'),Challenge=c('Lc','Mc','Hc'))
traitLabels  <- makeTraitLabels(trait=1,ntrait=length(traitLevels),levels=traitLevels)
StudentTrait <- setNames(strsplit(traitLabels,'[a-z]'),traitLabels)
#=====================================================
```
Modal States of Each Student Property:
```{r}
a <- studentBF %>% dplyr::select(probSchool,starts_with('BF')) %>% as.matrix
meanBF <- t(a[,1] %*% a[,-1])
# *** PROBLEMATIC FOR MAINTENANCE & REVISION: MUST RESET THESE IF ANY COLUMNS CHANGE IN studentBF ***
inm <- setNames(c(2,6,9,1,1,5,8,8,1,5,1,1,5,38),
                c('dependent','income','ethnicity','gender','pellgrant','fasfa',
                  'locale','region','defaultrate','earnings','admissions',
                  'completion','sat','discipline')) 
propNames <- mapply(function(i,si) rownames(meanBF)[seq(si-i+1,length.out=i)],inm,cumsum(inm))
domProps  <- sapply(propNames,function(nms) names(meanBF[nms,])[which.max(meanBF[nms,])])
show(domProps)
```
Create a few specific student profiles

(Check function 'utility' for valid entries for the student profile properties:
valid entries appear as the names() of vectors 'ethlist', 'inclist', etc.)
```{r}
students <- list(
  Black = list(
    dependent = TRUE,
    ethnicity = 'black',
    gender    = 'female',
    age       = 'le24',
    income    = 'gt30Kle110K',
    earnings  = 'gt110K',
    sat       = 'gt1000le1200',
    fasfa     = 'fsend_4',
    discipline= 'SocialSciences',
    region    = 'MidEast',
    locale    = 'CityLarge',
    traits    = StudentTrait$MrMvMbMc
  ),
  White = list(
    dependent = TRUE,
    ethnicity = 'white',
    gender    = 'female',
    age       = 'le24',
    income    = 'gt30Kle110K',
    earnings  = 'gt110K',
    sat       = 'gt1000le1200',
    fasfa     = 'fsend_4',
    discipline= 'SocialSciences',
    region    = 'MidEast',
    locale    = 'CityLarge',
    traits    = StudentTrait$MrMvMbMc
  ),
  Hispanic = list(
    dependent = TRUE,
    ethnicity = 'hispanic',
    gender    = 'female',
    age       = 'le24',
    income    = 'gt30Kle110K',
    earnings  = 'gt110K',
    sat       = 'gt1000le1200',
    fasfa     = 'fsend_4',
    discipline= 'SocialSciences',
    region    = 'MidEast',
    locale    = 'CityLarge',
    traits    = StudentTrait$MrMvMbMc
  ),
  Asian = list(
    dependent = TRUE,
    ethnicity = 'asian',
    gender    = 'female',
    age       = 'le24',
    income    = 'gt30Kle110K',
    earnings  = 'gt110K',
    sat       = 'gt1000le1200',
    fasfa     = 'fsend_4',
    discipline= 'SocialSciences',
    region    = 'MidEast',
    locale    = 'CityLarge',
    traits    = StudentTrait$MrMvMbMc
  ))

```
## Visualizing Model Parameters
Let's look at plots of the model parameters for a few specific student types:
```{r}

# Test getParameters:
plotBeta <- function(beta,stdtProf){
  trtstr <- paste(paste(c('Risk','Vision','Breadth','Challenge'),stdtProf$traits,sep=":"),collapse=', ')
  b <- data_frame(Trait=factor(rownames(beta),levels=rownames(beta)[order(beta[,1],decreasing=TRUE)]),Value=beta[,1])
  gplt <- b %>% 
    ggplot(aes(x=Trait,y=Value,fill=Trait)) 
  (gplt + 
    geom_bar(stat='identity',position='dodge') + 
    #theme(text=element_text(face='bold')) + 
    scale_y_continuous(lim=c(-1,1)/2) +
    ggtitle(sprintf("Utility Partworths for %s",trtstr))) %>% print
}
```
Note how the parameters change with the behavior traits of Risk, Vision, Breadth & Challenge.
```{r}
stdtProf <- students$White
beta1 <- getParameters(stdtProf,propertyMap)
plotBeta(beta1,stdtProf)

stdtProf <- students$White
stdtProf$traits <- StudentTrait$HrLvLbHc
beta2 <- getParameters(stdtProf,propertyMap)
plotBeta(beta2,stdtProf)

stdtProf <- students$White
stdtProf$traits <- StudentTrait$LrHvHbLc
beta3 <- getParameters(stdtProf,propertyMap)
plotBeta(beta3,stdtProf)
```

# Generating Personalized College Rankings
What follows is function `studentCaseStudy` the 'work horse' function to
provide the student with personalized college rankings:

```{r}
# Define a function to assess the suitability of colleges for a specific student's profile:
studentCaseStudy <- function(studentBF,stdtProf,propertyMap,ntop=20,signifLevel=0.15,verbose=FALSE,exact=FALSE){
  utilities <- studentBF %>% 
    dplyr::select(starts_with('BF')) %>%
    apply(1,utility,studentProfile=stdtProf,propertyMap=propertyMap) %>%
    setNames(studentBF$College)
  
  tmp <- studentBF %>% 
    dplyr::select(starts_with('BF')) %>%
    slice(1) %>% as.matrix %>%
    utility(studentProfile=stdtProf,propertyMap=propertyMap,dump.covariates=TRUE)
  
  wxBF <- studentBF %>% 
    dplyr::select(starts_with('BF')) %>%
    apply(1,utility,studentProfile=stdtProf,propertyMap=propertyMap,dump.covariates=TRUE) %>%
    t %>% as.data.frame %>% tbl_df %>%
    setNames(gsub('^[^:]+:','',colnames(tmp))) %>%
    mutate(unitID  = studentBF$unitID,
           College = studentBF$College,
           Utility = utilities) %>% 
    arrange(desc(Utility)) %>%
    dplyr::select(unitID,College,Utility,everything())
  
  labels <- sprintf("%d. %s",order(order(utilities,decreasing=TRUE)),names(utilities))
  
  stdtUtility <- data_frame(unitID  = studentBF$unitID, 
                            College = names(utilities),
                            Utility = utilities) %>%
    mutate(labels = factor(labels,levels=labels[order(Utility,decreasing=TRUE)]),
           expChoice = 10^Utility,
           ChoicePct = 100*expChoice/sum(expChoice)) %>%
    dplyr::select(-expChoice) %>%
    inner_join(wxBF %>% dplyr::select(-College,-Utility),by='unitID') %>% 
    arrange(desc(Utility)) #%T>% print
  
  topN <- stdtUtility %>% dplyr::select(-College) %>% top_n(ntop,Utility) 
  
  if(verbose) {
    topN %>% print
    traitStr <- paste(paste(c('Risk','Vision','Breadth','Challenge'),stdtProf$traits,sep=":"),collapse=', ')
    profText <- stdtProf[!grepl('beta|traits',names(stdtProf))]
    pltTitle <- paste(strwrap(paste(paste(names(profText),profText,sep=':'),collapse=', '),width = 80),collapse='\n')
    pltTitle <- paste0("Top ",ntop," Colleges for Profile:\n",pltTitle,'\n',traitStr)
    
    gplt0 <- topN %>% 
      mutate(labels = factor(labels,levels=rev(levels(labels))),
             labelY = pmin(0,min(Utility))) %>%
      ggplot(aes(x=labels,y=Utility,fill=labels))
    gplt <-  gplt0 +
      geom_bar(stat='identity',position='dodge') + coord_flip() + 
      geom_text(aes(x=labels,y=labelY,label=labels),size=6,hjust=0,vjust=0.5) +
      theme(text=element_text(face='bold',size=8),
            axis.text.y=element_blank(),
            title=element_text(size=12,face='bold'),
            legend.position = "none") +
      ggtitle(pltTitle) +
      labs(x=sprintf("Top %d Colleges",ntop)) 
    gplt %>% print
  }
  
  # Back-calculate the beta coefficients and see if can determine which
  # covariate contributed most strongly to the utility rankings.
  stdtBF <- stdtUtility %>% 
    inner_join(studentBF %>% dplyr::select(-College),by='unitID')
  axbf <- stdtBF %>% dplyr::select(one_of(gsub('^[^:]+:','',colnames(tmp)))) %>% as.matrix
  betacoeff <- solve(t(axbf) %*% axbf,t(axbf) %*% matrix(stdtUtility$Utility,ncol=1))
  
  if(verbose) show(betacoeff[betacoeff>1.0E-9,])
  
  stdtBF %<>%
    dplyr::select(c(4,3,5),one_of(names(betacoeff[abs(betacoeff)>1.0E-9,]))) 
  
  if(verbose) stdtBF %>% print(n=ntop,width=Inf)
  
  utlBFcor <- stdtBF %$% 
    lapply(.[-(1:3)],function(y) cor.test(Utility[1:ntop],y[1:ntop],method='spearman',exact=exact)) #%T>% print
  
  # This shows which covariates dictate the rankings by utility. If significant
  # negative correlation, then that covariate is important in the latter half of
  # the ranking (note that reverse ordering due to negative coefficients is
  # already accounted for).
  pvals <- sapply(utlBFcor,'[[','p.value')
  pvals <- pvals[!is.na(pvals)]
  
  signifCor <- utlBFcor[pvals <= pmax(signifLevel,min(pvals,na.rm=TRUE))] %$% {.[order(sapply(.,'[[','p.value'))]}
  
  if(verbose) signifCor %>% show
  
  results <- list(student=stdtProf,topN=topN,utilities=stdtUtility,BF=stdtBF,cor=utlBFcor,signifCor=signifCor,beta=betacoeff)
  invisible(results)
}
```
# Student Case Studies
Ideally, we'd build a nice GUI/app around this model so that a student can
enter his or her specific information and get a custom ranking of colleges.
Here, we'll just manually specify a student and assess the results:
```{r}
# Assess a specific student's case:
stdtProf <- students$Black
#stdtProf$gender <- 'male'
stdtProf$beta <- getParameters(stdtProf,propertyMap)
caseResult    <- studentCaseStudy(studentBF,stdtProf,propertyMap,verbose=TRUE,ntop=20)
```
# How About a Road-Trip to Your Top-20 Schools?!
I've copied idea and code snips from Tad Dallas's script on Kaggle.com:
```{r eval=TRUE, echo=TRUE, tidy=TRUE}
mortarBoardIcon <- makeIcon(
  iconUrl = "https://upload.wikimedia.org/wikipedia/commons/thumb/7/76/Fxemoji_u1F393.svg/512px-Fxemoji_u1F393.svg.png",
  iconWidth = 30, iconHeight = 30)

schools <- caseResult$topN %>%
  inner_join(student %>% 
               dplyr::select(c(unitID,College,city,state,zip,median_hh_inc_2005,lon,lat)), by='unitID') %>%
  mutate(info = paste0(College,'<br>',city,', ',state,' ',zip,
                       sprintf('<br> Rank: %d Utility: %.2f',seq_along(Utility),Utility)))
map <- leaflet(schools) %>% 
  setView(-93.65, 42.0285, zoom = 4) %>%
  addTiles() %>%
  addMarkers(~lon, ~lat, popup=~info,
             options = popupOptions(closeButton = TRUE),
             clusterOptions = markerClusterOptions(), 
             icon = mortarBoardIcon)
#print(map)
map

```


# Sensitivity Analysis

Finally, we'll do a brief set of sensitivity analyses illustrating how
the college rankings change with the student's demographics and behavioral
traits.  

We generate barplots of the utilities of the top 20 colleges for specific
student profiles. (You may want to just execute
`print(sensResults$grobList)` at the console after running the script
just to see individual top-10 school barplots in a legible format.)
```{r}
#=======================================
### SENSITIVITY ANALYSIS:
# Define a function for one-variable-at-a-time analysis of the student profile properties:
#if(FALSE){
# paropts <- par(no.readonly=TRUE)
sensitivity <- function(studentBF,stdtProf0,pmap,StudentTrait,propertyMap,maxcnt=Inf,plot.it=TRUE){
  #require(gridExtra)
  propnms <- names(pmap)
  cnt <- 0
  
  caseList <- list()
  grobList <- list()
  for(propnm in propnms){
    stdtProf <- stdtProf0
    # par(mfrow=c(2,ceiling(length(pmap[[propnm]])/2)))
    if(propnm == 'income') {
      if(stdtProf$dependent) {
        lvls <- names(pmap[[propnm]]$dependent)
      } else {
        lvls <- names(pmap[[propnm]]$independent)
      }
    } else {
      lvls <- names(pmap[[propnm]])
    }
    for(levelnm in lvls){
      stdtProf[[propnm]] <- levelnm
      stdtProf$beta <- getParameters(stdtProf,propertyMap)
      caseResult    <- studentCaseStudy(studentBF,stdtProf,propertyMap,ntop = 10,verbose = FALSE)
      
      profText <- c(stdtProf[propnm],stdtProf[!grepl(paste0('beta|traits|',propnm),names(stdtProf))])
      pltTitle <- paste(strwrap(paste(paste(names(profText),profText,sep=':'),collapse=', '),width = 80),collapse='\n')
      traitStr <- paste(paste(c('Risk','Vision','Breadth','Challenge'),stdtProf$traits,sep=":"),collapse=', ')
      pltTitle <- paste0("Top 10 Colleges for Profile:\n",pltTitle,'\n',traitStr)
      
      nms <- names(caseList)
      caseList <- setNames(c(caseList,list(caseResult)), c(nms,pltTitle))
      
      #if(plot.it){
      gplt0 <- caseResult$topN %>% 
        mutate(labels = factor(labels,levels=rev(levels(labels))),
               labelY = pmin(0,min(Utility))) %>%
        ggplot(aes(x=labels,y=Utility,fill=labels))
      gplt <-  gplt0 +
        geom_bar(stat='identity',position='dodge') + coord_flip() + 
        geom_text(aes(x=labels,y=labelY,label=labels),size=6,hjust=0,vjust=0.5) +
        theme(text=element_text(face='bold',size=8),
              axis.text.y=element_blank(),
              title=element_text(size=12,face='bold'),
              legend.position = "none") +
        ggtitle(pltTitle) +
        labs(x="Top 10 Colleges") 
      #gplt %>% print
      grobList <- c(grobList,list(gplt))
      #}
      
      cnt <- cnt + 1
      if(cnt>=maxcnt) break
    }
    if(cnt>=maxcnt) break
  }
  # par(paropts)
  
  ### UNFORTUNATELY, SEEMS TO LOSE CONTROL OF TEXT SIZE WHEN ggplots ARE LAID IN
  ### GRID USING PACKAGE 'gridExtra'.... HELP!!!
  #gmulti <- marrangeGrob(grobList,nrow=2,ncol=4)
  if(plot.it) do.call(grid.arrange,grobList)
  return(list(caseList=caseList,grobList=grobList))
}
```
## Sensitivity Analysis #1:
Perform sensitivity analysis for a stellar female student interested in 
Engineering and a Large City setting, with a trait profile of Risk=High, 
Vision=Low, Breadth=High, Challenge=High. We'll sweep over regions. 

Note that because Breadth is High, she won't stick to strictly engineering
schools. And because Vision is Low, she stresses region and locale more
strongly than future earnings.
```{r}
stdtProf0 <- students$White
stdtProf0$gender <- 'female'
stdtProf0$discipline <- 'Engineering'
stdtProf0$sat <- 'gt1400'
stdtProf0$traits <- StudentTrait$HrLvHbHc # Risk=High, Vision=Low, Breadth=High, Challenge=High
pmap <- propertyMap[c('region','ethnicity')]
#pmap$ethnicity <- pmap$ethnicity[1:4]
maxcnt <- length(pmap[[1]]) # Only flip ethnicity
sensResults <- sensitivity(studentBF,stdtProf0,pmap,StudentTrait,propertyMap,maxcnt = maxcnt, plot.it = FALSE)
print(sensResults$grobList)
```

## Sensitivity Analysis #2: 
Perform sensitivity analysis for a stellar male student interested in
Engineering and a Large City setting, with a trait profile of Risk=Low,
Vision=Low, Breadth=High, Challenge=High. We'll sweep over ethnicity; then
for fixed ethnicity=White, we'll sweep over region.

Note that because Risk is Low, he'll strongly stress region & seek student
populations that look a little more like himself.

And in the case of Region = New England, the student's top-10 even has schools
with negative utility!
```{r}
stdtProf0 <- students$White
stdtProf0$gender <- 'male'
stdtProf0$discipline <- 'Engineering'
stdtProf0$sat <- 'gt1400'
stdtProf0$traits <- StudentTrait$LrLvHbHc # Risk=Low, Vision=Low, Breadth=High, Challenge=High
pmap <- propertyMap[c('ethnicity','region')]
pmap$ethnicity <- pmap$ethnicity[1:4]
pmap$region    <- pmap$region[c('NewEngland','Southeast','GreatLakes','FarWest')]
maxcnt <- Inf
sensResults <- sensitivity(studentBF,stdtProf0,pmap,StudentTrait,propertyMap,maxcnt = maxcnt, plot.it = FALSE)
print(sensResults$grobList)
```

## Sensitivity Analysis #3: 
Now, let's see what happens when that same stellar male student shows up with a trait
profile with more risk and has a longer-term vision:
```{r}
stdtProf0 <- students$White
stdtProf0$gender <- 'male'
stdtProf0$discipline <- 'Engineering'
stdtProf0$sat <- 'gt1400'
stdtProf0$traits <- StudentTrait$HrHvHbHc # Risk=High, Vision=High, Breadth=High, Challenge=High
pmap <- propertyMap[c('ethnicity','region')]
pmap$ethnicity <- pmap$ethnicity[1:4]
pmap$region    <- pmap$region[c('NewEngland','Southeast','GreatLakes','FarWest')]
maxcnt <- Inf
sensResults <- sensitivity(studentBF,stdtProf0,pmap,StudentTrait,propertyMap,maxcnt = maxcnt, plot.it = FALSE)
print(sensResults$grobList)
#}
#------------------------------------------------------------
```

## Comments on Sensitivity Analyses
As you can see, the student's value assessment, in terms of beta parameters
(partworths), as determined by his or her behavioral traits and preferences
strongly affects the top 10 high-utility schools for that student.

Experiment with different student profiles, such as lower SAT's, and interest
in alternative settings (locales other than Large City). The results are
fascinating!

Even better, alter the beta parameters directly yourself in `stdtProf0$beta`....

# Next Steps
Please, feel free to extend this model and tool. It'd be great to slap a
quick Shiny interface on it so that a student can input his or her own
demographics and criteria and quickly generate lists and charts detailing the
student's best college options.

Being posed in a probabilistic discrete choice modeling framework, this model
represents lots of potential. It can, for example, be presented with
individual student data -- imagine the entire dataset of applicants to a
university plus the indication of which applicants were accepted or not --
and then, using tools such as the package `rstan`, we could estimate
data-based model parameters beta suitable for identifying candidates for a
specific college.  This essentially reverse engineers the college's
admissions criteria.

It also offers a principled way of identifying which specific college is or
is not a good choice for a student. 

Even in its simplistic form shown here -- which didn't implement even some of
the computed college properties -- it's quite entertaining to see how varied
the rankings are given specific students -- e.g., a top student (SAT>1400)
with a strong regional preference makes quite interesting choices as the
region of interest changes. Try it out!

Any feedback would be appreciated.

Thanks,

-Michael L. Thompson 

[(my info at LinkedIn)](<https://www.linkedin.com/in/mlthomps>)
